### 2015-02-24 Paper.tex
initialised

\documentclass[conference]{IEEEtran}

\usepackage{graphicx}

\graphicspath{ {images/} }

\usepackage{listings}
\usepackage{color}

\definecolor{dkgreen}{rgb}{0,0.6,0}
\definecolor{gray}{rgb}{0.5,0.5,0.5}
\definecolor{mauve}{rgb}{0.58,0,0.82}

\lstset{
	language=Java,
	aboveskip=3mm,
	belowskip=3mm,
	showstringspaces=false,
	columns=flexible,
	basicstyle={\small\ttfamily},
	numbers=none,
	numberstyle=\tiny\color{gray},
	keywordstyle=\color{blue},
	commentstyle=\color{dkgreen},
	stringstyle=\color{mauve},
	breaklines=true,
	breakatwhitespace=true,
	tabsize=3
}

\begin{document}
	\title{Hadoop Image Processing Framework}
	\author{\IEEEauthorblockN{Sridhar Vemula}
		\IEEEauthorblockA{Computer Science Department\\
			Oklahoma State University\\
			Stillwater, Oklahoma \\
			Email: sridhar.vemula@okstate.edu}
		\and
		\IEEEauthorblockN{Christopher Crick}
		\IEEEauthorblockA{Computer Science Department\\
			Oklahoma State University\\
			Stillwater, Oklahoma \\
			Email: chriscrick@cs.okstate.edu}
	}
	\maketitle
	\begin{abstract}
		With the rapid growth of social media, the number of images being uploaded to the internet is exploding. Massive quantities of images are shared through multi-platform services such as Snapchat, Instagram, Facebook and WhatsApp; recent studies estimate that over 1.8 billion photos are uploaded every day.  However, for the most part, applications that make use of this vast data have yet to emerge. Most current image processing applications, designed for small-scale, local computation, do not scale well to web-sized problems with their large requirements for computational resources and storage.  The emergence of processing frameworks such as the Hadoop MapReduce\cite{Hadoop2009} platform addresses the problem of providing a system for computationally intensive data processing and distributed storage. However, to learn the technical complexities of developing useful applications using Hadoop requires a large investment of time and experience on the part of the developer.  As such, the pool of
		researchers and programmers with the varied skills to develop applications that can use large sets of images has been limited. To address this we have developed the Hadoop Image Processing Framework, which provides a Hadoop-based library to support large-scale image processing. The main aim of the framework is to allow developers of image processing applications to leverage the Hadoop MapReduce framework without having to master its technical details and introduce an additional source of complexity and error into their programs.
	\end{abstract}

	\IEEEpeerreviewmaketitle

	\section{Introduction}
	With the spread of social media in recent years a large amount of image data is accumulated. Processing this large image data is limited to single computer which lack in computational power and storage ability. These tasks typically be performed on a distributed system by dividing the task into sever sub tasks. The ability to parallelize tasks allows for scalable, efficient, execution of resource-intensive applications. Hadoop MapReduce framework provides a platform for such tasks.

	Also, when considering operations such as face detection, image classification\cite{Li2009} and other types of processing on images, there are limits on what can be done to improve performance of single computers to make them able to process large-scale information. Therefore, the advantages of parallel distributed processing of a large image dataset by using the computational resources of cloud computing environment can be considered. In addition, if computational resources can be secured easily and relatively inexpensively, then cloud computing in suitable for handling large size image data sets at very low cost and increased performance. Hadoop, as a system for processing large numbers of images by parallel and distributed computing seems promising. In fact, Hadoop is in use all over the world. Studies using Hadoop have been performed to treat one file as a text data file or multiple files as a single file unit, such as for the analysis of large volumes of DNA sequence data, converting the data of a large number of still images to PDF format, and carrying out feature selection/extraction in astronomy. These examples demonstrate the usefulness of this system, which is due to its having the ability to run multiple processes in parallel for load balancing and task management.

	Most of the image processing applications, that use Hadoop MapReduce framework require a staggering learning curve and high complexity. The overhead required to implement such applications is cumbersome to users. Hadoop Image Processing Framework hides the highly technical deatails of Hadoop system and provides users with a feel of an normal image processing library giving the users the access to advanced resources of distributed system. Our framework mainly focuses around giving users an easy way to access the image data allowing for easy and flexible use for large scale image applications

	Hadoop Image Processing Framework is largely a software design project, with a goal of hiding Hadoop's functionality and provide users with a library to effectively use Hadoops MapReduce system for large scale image processing. We believe that the ease of use and java oriented semantics will make the process of creating large scale image experiments further more easier. This framework is an excellent tool for Hadoop novice users and image processing researchers because it allows development of large scale image processing applications.

	In the following section  we will describe prior work in this area. Next we describe the overview for the Hadoop Image Processing Framework including the Downloader, Processor and Extractor statges. Additionally, we describe our approach of distributing tasks for MapReduce. Finally we demonstrate the potential of this framework with a few examples of image processing.

	\section{Prior Work}
	With rapid increase of usage of online photo storage and social media like facebook, flickr and picasa the amount of image data available is larger than ever before. The amount of image data is growing rapidly every day. According to Buzzfeed for every minute 27800 photos are uploaded to Instagram. As per facebook, they get 208,300 photos uploaded every minute\cite{Horaczek2013}. This alone provides a plethora of image that can scale up to billions of images. It is these reasons that motivate the need for research with image processing applications that take advantage of large sets of images

	A brief overview of previous work sharing the themes of image processing and distributed computation is presented below:

	Web-Scale Computer Vision using MapReduce for Multimedia Data Mining \cite{White2010} present a case study of classifying and clustering billions of regular images using MapReduce. A way of pre-processing images for use in a sliding-window approach for object recognition is described. Distributed High Performance Video Processing in the cloud, outlines some of the limitations of the MapReduce model when dealing with high-speed video encoding, namely its dependence on the NameNode as a single point of failure, and lack of possibility for generalization in order to suit the issue at hand. An alternate optimized implementation is proposed for providing cloud-based IaaS (Infrastructure as a Service) Solution. In Parallel K-Means clustering of Remote Sensing Images Based on MapReduce \cite{Lv2010} describe using the k-means algorithm in conjunction with MapReduce and satellite/aerophoto images in order to find different elements based on their color. Using Transaction Based Parallel computing to solve image processing and computational physics problems by Harold describes the use of distributed computing with two examples video processing and subsurface transport. There was no information provided on how image processing parts of the examples.

	Case Study of Scientific Data Processing on a Cloud Using Hadoop \cite{Zhang2010} present methods used for processing sequences of microscope images of live cells. The images are relatively small 512x512 16-bit pixels, stored in 90 MB folders there were some issues regarding to fitting into Hadoop DFS blocks with were solved by custom InputFormat, InputSplit and Record Reader classes. A scalable image processing framework for gigapixel Mars and other celestial body images by Mark Powell describes the way NASA handles image processing of celestial images captured by the Mars Orbiter and rovers. Clear and concise descriptions are provided about the segmentation of gigapixel images into tiles, how the tiles are processed and how image processing framework handles scaling and works with the distributed processing. Ultra-fast processing of gigapixel tissue microarray images using high performance computing \cite{Wang2011} about the speeding up the analysis of Tissue MicroArray images by substituting human expert analysis for automated processing algorithms. While the images were measured to be in gigapixels, the content is easily segmented and there was no need to focus on being able to analyze all of image at once. The work was all done on a specially build grip high performance	computing platform on Hadoop framework.

	Terabyte size image computations on Hadoop cluster \cite{Bajcsy2013} platforms present a characterization of four basic terabyte size image computations on a Hadoop Cluster in terms of their relative efficiency according to the modified Amdahls law. The work is motivated by the fact that there is a lack of 6 standard benchmarks and stress tests for big image processing operations on Hadoop framework. Terabyte-scale image search\cite{Moise2013} outlines about querying thousands of images in one run using Hadoop MapReduce framework using the MapReduced based eCP Algorithm. The experiment does an image search on 110 million images collected from web on Grid 5000 Platform. The results are evaluated in order to understand the best practices that can tune Hadoop MapReduce Performance for image search.

	While the above shows that there has been lot of work in this area, providing a framework to ease the process develop large-scale image processing applications. The question remains how the performance for Hadoop can be increased for large scale image processing tasks on small files.

\section{Methodology}
Hadoop Image processing Framework is created to provide users a tool to develop a large scale image processing applications with ease.

The main goals of Hadoop Image Processing framework:
\begin{itemize}
	\item Provide an open source framework over Hadoop MapReduce for developing large scale image applications 
	\item Flexibility to store images in various Hadoop File Formats
	\item Present users with an intuitive interface for image based operations which are highly parallelized and balanced and hide the high technical details of hadoop MapReduce Framework
	\item Allow interoperability between various image processing libraries
\end{itemize}

\subsection{Downloading and storing image data}
Hadoop's distributed file system(HDFS) to store files in various nodes throughout the cluster. Storing small files\cite{White2009} is a big problem in Hadoop. A small file is one which is significantly smaller than HDFS block size. As the large image dataset can have many number of small files and the problem is HDFS cant handle lots of files. The problem can be solved by providing some kind of container to group the files in some way. Hadoop offers a few options:
\begin{itemize}
	\item HAR File
	\item Sequence File
	\item Map File
\end{itemize}

The Downloader Module of the framework does the following work

\textbf{Step 1: Input a URL List.} Initially user inputs a file containing url's of images to download. The input list should be a text file with one image url per line. The list can be generated by hand, database or a search. The framework provides a extendable ripper module to extract urls from flickr, google and databases. In addition to the list the user will input the type of image bundle to be generated(e.g. har, sequence or map). According to the input, the urls are divided across the nodes for maximum efficiency and parallelism to download the images. Each node map task will generate few image bundles as per the input type containing all the of the images it downloaded. In the reduce phase, the Reducer will merge these image bundles  into a large image bundle.

\textbf{Step 2: Split URLs across nodes.}
From the input, file containing list of image urls and the type of file to be generated, we equally distribute the task of downloading images across the all the nodes in the cluster. The nodes are efficiently managed so that no memory overflow can occur even for terabytes of image downloaded by a single map task. This allows maximum parallelization for downloading. Image urls are distributed to all nodes equally and different map tasks begin downloading the image set it is responsible for.

\begin{figure}[h]
	\centering
	\includegraphics[width=0.45\textwidth]{down-map}
	\caption{Individual Map task of Downloader Module}
	\label{fig:down-map}
\end{figure}

\textbf{Step 3: Download image data from URLs.}
For every url retrieved in the map task, a connection is established as per transfer protocal (e.g. ftp, http, https etc.). Once connected, we check the file type and if its a valid image we get an InputStream to the connection. From this InputStream, we generate a new HImage object and add the image to the image bundle. The HImage class holds the image data and helps user to play with image and image header data. HImage class also provides an interoperability between various image data types (e.g. BufferedImage, Mat etc.,).

\textbf{ Step 4: Store images in an image bundle. } 
Once HImage object is received, we can add the image to the image bundle simply by passing the HImage object to the appendImage method. Each map task generates a few or many image bundles depending on the image list. In the reduce phase, all these image bundles are merged into one large bundle.

\begin{figure}[h]
	\centering
	\includegraphics[width=0.45\textwidth]{down-node}
	\caption{Single Node running the Downloader Module handled by the framework and transparent to user}
	\label{fig:down-node}
\end{figure}

\subsection{Processing image bundle using MapReduce}
Hadoop MapReduce programs handle input and output data very efficiently, but the main problem is representing the image data. For instance, to distribute images across map nodes would require to pass images as string, then decode image into a specified format to access pixel information which is not only inefficient but also inconvenient. To overcome this problem we want to able to represent the images as many different formats to increase flexibility. The framework focuses on bringing familiar data types directly to user.

As distribution is important in MapReduce, we want to process the image in the same machine where the bundle block resides. In general, the user needs to create an InputFormat and Record Reader classes to specify the MapReduce job and distribute the input among nodes. This task is a bit cumbersome. In order to make the framework more easier we took care of the InputFormat and RecordReaders.

Images are distributed as various image data types and users can immediately have access to pixel values. As we attempt to obtain the pixel values we lose the valuable image header data(e.g. jpeg exif data, IHDR\cite{David03} Image Header). The framework, holds the image data in a special HImageHeader data type before converting the image bytes into pixel values. After working on the pixel data, it was made sure that image header is appended to the processed pixels. Even though it adds an storage overhead, image header data is preserved even after processing.

The working of Processor module of the framework is described below:

\textbf{Step 1: Input the Algorithm.} We assume that the user writes an algorithm which extends a Generic Algorithm class. This class is passed as an argument to the processor module. The framework starts a MapReduce Job with the algorithm as an input. The GenericAlgorithm holds an HImage variable, this helps user to write an algorithm on the single image which iterates over the whole image bundle. In addition to the algorithm the user should input the image bundle file that needs to be processed. According to the input the image bundle is divided across nodes as individual map task. Each Map task will process the image as of the algorithm and appends them to the output image bundle. In the reduce phase, the Reducer will merge these image bundles into a large image bundle.

\textbf{Step 2: Split Imagebundle across nodes.} The input image bundle is stored in HDFS as blocks. In order to obtain maximum throughput it was made sure the map task is run in the same the block where it resides, using custom input format and record reader classes. This allows maximum parallelization without a problem of transferring data across nodes. The image bundle now executes different map tasks for the image data it is responsible for.

\begin{figure}[h]
	\centering
	\includegraphics[width=0.45\textwidth]{pro-map}
	\caption{Individual Map task of Processor}
	\label{fig:pro-map}
\end{figure}

\textbf{Step 3: Process individual image.} For every HImage retrieved in the map task, the algorithm is run on the HImage which is given as the input to the processor module. The HImage retrieves the image as data type(e.g. BufferedImage, Mat image) in the algorithm. Once the image data type is retrieved processing takes place. After processing, the image header data of the original image is appended to the processed image, thereby preserving image header data. The processed image is appended to the temporary bundle generated by the map task.

\textbf{Step 4: Store processed images in an image Bundle.} Every map task generate a image bundle after the map task is completed. Once the map phase is completed there are many bundles. In the reduce phase, all these temporary image bundles are merged into a single large file which contains all the processed images.

\begin{figure}[h]
	\centering
	\includegraphics[width=0.45\textwidth]{pro-node}
	\caption{Single Node running the Processor Module handled by the framework and transparent to user}
	\label{fig:pro-node}
\end{figure}

\subsection{Extracting image bundle using MapReduce}
The framework creates and processes image bundles but there should be a way to extract these images and make them viewable. Generally, Hadoop extracts images from the image bundle iteratively using a single node which is not efficient. To make the process distributed among different nodes we designed an extractor module which can extract images in parallel in all the nodes.

Distribution plays a key role in MapReduce, we want to make use of the nodes effectively to make the process of extraction efficient. In general, the user needs to create custom InputFormat and RecordReader classes to specify the MapReduce job and distribute the input among nodes. In order to over come this the framework takes care of these Custom InputFormat and RecordReader to make working with hadoop more easier.

There is often a confusion to where the images need to be extracted. The images can be extracted to the local file system of the node or hadoop file system. The framework overcomes the problem by providing a flexibility to choose.

The working of Extractor module is explained below.

\textbf{Step 1: Input the image bundle to be extracted.} The image bundle that needs to extracted is passed as a parameter to the Extractor module. Along with it the user can send an optional parameter to where the images need to be extracted (defaults to local file system). After the input the image bundle is distributed across nodes as individual map task. Each map task will extract the image in to the filesystem.

\begin{figure}[h]
	\centering
	\includegraphics[width=0.45\textwidth]{ext-map}
	\caption{Individual Map Task of Extractor Module}
	\label{fig:ext-map}
\end{figure}

\textbf{Step 2: Split Image bundle across nodes}
The input image bundle split across nodes using the custom input format and record reader classes for maximum throughput. Once a Map task starts a HImage object is retrieved.

\textbf{Step 3: Extract individual image}
The obtained HImage object is retrieved as image bytes and the image bytes are stored into the file based on its image type. The Extractor module doesn't have a reduce phase.

\begin{figure}[h]
	\centering
	\includegraphics[width=0.45\textwidth]{ext-node}
	\caption{Single node running the Extractor Module handled by the framework and transparent to user}
	\label{fig:ext-node}
\end{figure}

\section{Experiments and Observations}

We describe two image applications using Hadoop Image Processing framework. These applications enable user perform large scale image processing operations with ease. These applications are complex and inefficient one existing platforms, yet simple to implement using the framework

\subsection{Canny Edge Detection}

Canny Edge Detection\cite{Canny86} is an multi-stage algorithm to detect a wide range of edges in images. Edge detection, especially step edge detection is an important technique to extract useful structural information from different vision objects and dramatically reduce the amount of data to be processed. Among the edge detection methods developed so far, canny edge detection algorithm is one of the most strictly defined methods that provides good and reliable detection.

\begin{figure}[h]
	\centering
	\includegraphics[width=0.45\textwidth]{input-canny}
	\caption{A Sample Image from downloaded image bundle}
	\label{fig:input-canny}
\end{figure}

\begin{figure}[h]
	\centering
	\includegraphics[width=0.45\textwidth]{output-canny}
	\caption{Processed image using canny edge detection}
	\label{fig:output-canny}
\end{figure}

\subsubsection{Characterstics of Image Dataset, Hardware and Software}
We computed edge detection on a dataset which is about 1TB size. The data set is made from the flickr search based on a keyword flight. The data set is extracted using the a FlickrRipper in the framework which implements Ripper Interface. The Ripper interface can be implemented for any custom Rippers to get the image URLs. The example data set is composed of 220000 images at 4.76 MB/image equals to 1 TB. We ran the edge detection on the cluster with 7 nodes having 2 to 4 virtual processors with 4 to 8 GB RAM. Each virtual processor clocks at a speed of 3.2 Ghz. The algorithm was implemented independently and used on single desktop, Hadoop cluster without framework and Hadoop cluster with framework.

\subsubsection{Observations}

The desktop implementation of the Algorithm is a java program for processing a small image dataset. As soon as the data set is processed the program halts. The desktop implementation starts as a single thread. One image is processed every time. As soon as the image is processed, it starts processing a new image. When the data set size is low, the program runs smooth with best performance. As data set size increases to Terabytes the system can't handle such huge amounts of data. To handle such large amounts of data we use the distributed computing using hadoop.

\begin{figure}[h]
	\centering
	\includegraphics[width=0.45\textwidth]{module-chart}
	\caption{Module Chart}
	\label{fig:module-chart}
\end{figure}

The Algorithm is implemented on a hadoop cluster without using framework. It overcomes the problem of handling large datasets but suffers with a problem of code complexity. The user needs to write a huge amount of code and understand the technical details of hadoop. Writing such huge amounts of code including the custom InputFormats and RecordReaders is cumbersome. Fig.\ref{fig:module-chart} compares the actual lines of code user needs to write to run an Image Processing application in hadoop cluster which makes an average user life harder to use the hadoop features.

The Algorithm is finally implemented on a hadoop cluster using the framework. It overcomes many problems handling large datasets, code complexity, interoperability between image datatypes etc., The process of creating CustomInputFormat and RecordReader is automated as per the input. User can tune the settings using different getter and setter method to get the desired output. The framework is so transparent that the user needs to input few lines of code to process a large image data set. It hides all the MapReduce Technical details and runs the code highly parallelized. The key differences among the three platform-specific implementation lie in computation elasticity and code complexity associated with each platform.

\begin{table}[h]
	\begin{tabular}{|l|l|l|l|}
		\hline
		\multicolumn{1}{|c|}{\textbf{\begin{tabular}[c]{@{}c@{}}Computa-\\ tional\\ Platform\end{tabular}}} & \multicolumn{1}{c|}{\textbf{\begin{tabular}[c]{@{}c@{}}Computational \\ Elasticity\end{tabular}}}      & \multicolumn{1}{c|}{\textbf{\begin{tabular}[c]{@{}c@{}}Code \\ Complexity\end{tabular}}}                                                                                                                                                  & \multicolumn{1}{c|}{\textbf{\begin{tabular}[c]{@{}c@{}}Data \\ Locality\end{tabular}}}                                                                           \\ \hline
		Desktop                                                                                             & \begin{tabular}[c]{@{}l@{}}Low : limited by \\ RAM and CPU of \\ the executing\\ computer\end{tabular} & \begin{tabular}[c]{@{}l@{}}Normal : Complexity \\ lies in writing the \\ algorithm\end{tabular}                                                                                                                                           & \begin{tabular}[c]{@{}l@{}}All data are \\ on local disk\end{tabular}                                                                                            \\ \hline
		\begin{tabular}[c]{@{}l@{}}Hadoop:\\ without \\ using \\ framework\end{tabular}                     & \begin{tabular}[c]{@{}l@{}}High: Nodes can\\  be requested \\ as needed\end{tabular}                   & \begin{tabular}[c]{@{}l@{}}Heavy : Complexity \\ lies inwriting every \\ module in framework \\ including custom \\ Input Formats and \\ RecordReaders.User \\ should have high \\ technical knowledge on \\ hadoop working.\end{tabular} & \begin{tabular}[c]{@{}l@{}}All data resides\\ in HDFS. Custom\\ InputFormats \\ needsto be \\ created to launch\\ computations \\ where thedata are\end{tabular} \\ \hline
		\begin{tabular}[c]{@{}l@{}}Hadoop:\\ using \\ framework\end{tabular}                                & \begin{tabular}[c]{@{}l@{}}High: Nodes can \\ be requested \\ as needed\end{tabular}                   & \begin{tabular}[c]{@{}l@{}}Normal : Complexity \\ lies in writing \\ algorithm and is \\ equivalent to writing \\ in Desktop platform\end{tabular}                                                                                        & \begin{tabular}[c]{@{}l@{}}All data resides \\ in HDFS. Highly \\ parallelized.\end{tabular}                                                                     \\ \hline
	\end{tabular}
\end{table}

\iffalse
\begin{table}[h]
	\centering
	\begin{tabular}{|c|c|c|}
		\hline
		\textbf{Module} & \textbf{\begin{tabular}[c]{@{}c@{}}SLOC \\ using framework\end{tabular}}                                                          & \textbf{\begin{tabular}[c]{@{}c@{}}SLOC \\ without framework\end{tabular}}                                                                                    \\ \hline
		Downloader      & 3 - 5 lines                                                                                                                       & \begin{tabular}[c]{@{}c@{}}300 - 1000 lines.\\ Custom Input Formats\\ and RecordReaders are\\ to be implemented\end{tabular}                                  \\ \hline
		Processor       & 3 - 5 lines                                                                                                                       & \begin{tabular}[c]{@{}c@{}}400 - 1000 lines\\ Many overheads. \\ Converting to image \\ data types, Mapper \\ locality should be taken\\ care of\end{tabular} \\ \hline
		Extractor       & 2 - 8 lines                                                                                                                       & \begin{tabular}[c]{@{}c@{}}200 - 1000 lines\\ Depends on the extraction\\ process.\end{tabular}                                                               \\ \hline
		Algorithm       & \begin{tabular}[c]{@{}c@{}}200 - 300 lines\\ adds a few lines overhead\\ to automate the process\\ for MapReduce job\end{tabular} & 200 - 250 lines                                                                                                                                               \\ \hline
	\end{tabular}
\end{table}
\fi

Hadoop Image processing framework provides transparency over two levels. The first level of transparency allows user to use the predefined MapReduce modules Downloader, Processor and Extractor for processing the image modules. The second of level of transparency allows the users to write custom map reduce programs as the framework is built in an hierarchal manner.

The great thing about Hadoop Image Processing framework is its built in such a way that using it is really simple. The framework strictly adheres java file writing and reading techniques and extends it to a new level of BundleFile Writing and Reading using hadoop framework. This makes user to work with hadoop image processing more flexible and easier. The main aim of the framework is to provide an abstractness in such a way that programming on a single computer is equivalent to programming on a hadoop cluster.  Listing 1. provides the similarities in code writing in java and image processing framework. Making framework similar to java helps understanding the framework more easier and reduces complexity.

\begin{figure}[h]
	\centering
	\includegraphics[width=0.45\textwidth]{files-chart}
	\caption{Comparing File Writers and readers without and with using framework }
	\label{fig:files-chart}
\end{figure}

\begin{lstlisting}[caption = Comparing File Writer instance in java and SequenceBundleWriter instance in Hadoop Image Processing framework ]

//Sample File Writing in Java

File file = new File("abc.txt");
FileWriter fw=new FileWriter(file);
fw.append(val)
fw.close();

//Sample BundleFile Writing in Hadoop using Framework

BundleFile file = new BundleFile("abc.seq");
SequenceBundleWriter sbw=new SequenceBundleWriter(file);
sbw.append(himage);
sbw.close();

\end{lstlisting}

\section{Conclusion}
This paper has decribed our framework fo image processing for large scale image processing applications on the MapReduce framework. The framework is designed to abstract the technical details of Hadoop's powerful MapReduce framework and provide a way for users who want to process large image datasets. We provided a how the images can be stored in different hadoop file format and efficient access within Map Reduce pipeline and simple modules to make the framework more easier. By providing interoperability between different image data types we give the user to process the images using different open-source image processing libraries. Finally, we provide a way to overcome the loss of image headers after processing which are most useful for image processing and vision applications.

With these features, the framework brings a new level of transparency and simplicity for creating large-scale image processing applications over Hadoop's MapReduce framework. The paper describes to example applications built using the framework to demonstrate the power of the framework. We hope that this level of abstractness to use Hadoop MapReduce framework will enhance the ability to create large-scale image processing applications with ease.

\section{Future Work}
In the near future we hope to extend the framework into a full-fledged multimedia processing framework. We would like to improve framework such that audio and video processing over hadoop is done with ease. We also intend to add a CUDA module for using the graphics card for processing the images. Finally, we intend to wrap our system into an highly parallelized open-source hadoop multimedia processing framework and perhaps provide a sample web-based graphic user interface with all image processing applications.

	\bibliography{references}
	\bibliographystyle{IEEEtran}

\end{document}

### 2015-03-12 Paper.tex
Rewritten introduction

@@
 However, to learn the technical complexities of developing useful applications using Hadoop requires a large investment of time and experience on the part of the developer.
-As such, the pool of
-		researchers and programmers with the varied skills to develop applications that can use large sets of images has been limited.
+As such, the pool of researchers and programmers with the varied skills to develop applications that can use large sets of images has been limited.
 To address this we have developed the Hadoop Image Processing Framework, which provides a Hadoop-based library to support large-scale image processing.
@@
 	\section{Introduction}
-	With the spread of social media in recent years a large amount of image data is accumulated.
-Processing this large image data is limited to single computer which lack in computational power and storage ability.
-These tasks typically be performed on a distributed system by dividing the task into sever sub tasks.
-The ability to parallelize tasks allows for scalable, efficient, execution of resource-intensive applications.
-Hadoop MapReduce framework provides a platform for such tasks.
+	With the spread of social media in recent years, a large amount of image data has been accumulating.
+When processing this massive data resource has been limited to single computers, computational power and storage ability quickly become bottlenecks.
+Alternately, processing tasks can typically be performed on a distributed system by dividing the task into several subtasks.
+The ability to parallelize tasks allows for scalable, efficient execution of resource-intensive applications.
+The Hadoop MapReduce framework provides a platform for such tasks.
 
-	Also, when considering operations such as face detection, image classification\cite{Li2009} and other types of processing on images, there are limits on what can be done to improve performance of single computers to make them able to process large-scale information.
-Therefore, the advantages of parallel distributed processing of a large image dataset by using the computational resources of cloud computing environment can be considered.
-In addition, if computational resources can be secured easily and relatively inexpensively, then cloud computing in suitable for handling large size image data sets at very low cost and increased performance.
-Hadoop, as a system for processing large numbers of images by parallel and distributed computing seems promising.
+	When considering operations such as face detection, image classification\cite{Li2009} and other types of processing on images, there are limits on what can be done to improve performance of single computers to make them able to process information at the scale of social media.
+Therefore, the advantages of parallel distributed processing of a large image dataset by using the computational resources of a cloud computing environment should be considered.
+In addition, if computational resources can be secured easily and relatively inexpensively, then cloud computing is suitable for handling large image data sets at very low cost and increased performance.
+Hadoop, as a system for processing large numbers of images by parallel and distributed computing, seems promising.
 In fact, Hadoop is in use all over the world.
-Studies using Hadoop have been performed to treat one file as a text data file or multiple files as a single file unit, such as for the analysis of large volumes of DNA sequence data, converting the data of a large number of still images to PDF format, and carrying out feature selection/extraction in astronomy.
-These examples demonstrate the usefulness of this system, which is due to its having the ability to run multiple processes in parallel for load balancing and task management.
+Studies using Hadoop have been performed, dealing with text data files, analyzing large volumes of DNA sequence data, converting the data of a large number of still images to PDF format, and carrying out feature selection/extraction in astronomy.
+%Cite the papers in which these examples appear
+These examples demonstrate the usefulness of the Hadoop system, which can run multiple processes in parallel for load balancing and task management.
 
-	Most of the image processing applications, that use Hadoop MapReduce framework require a staggering learning curve and high complexity.
-The overhead required to implement such applications is cumbersome to users.
-Hadoop Image Processing Framework hides the highly technical deatails of Hadoop system and provides users with a feel of an normal image processing library giving the users the access to advanced resources of distributed system.
-Our framework mainly focuses around giving users an easy way to access the image data allowing for easy and flexible use for large scale image applications
+	Most of the image processing applications that use the Hadoop MapReduce framework are highly complex and impose a staggering learning curve.
+The overhead, in programmer time and expertise, required to implement such applications is cumbersome.
+To address this, we present the Hadoop Image Processing Framework, which hides the highly technical details of the Hadoop system and allows programmers who can implement image processing algorithms but who have no particular expertise in distributed systems to nevertheless leverage the advanced resources of a distributed, cloud-oriented Hadoop system.
+Our framework provides users with easy access to large-scale image data, smoothly enabling rapid prototyping and flexible application of complex image processing algorithms to very large, distributed image databases.
 
-	Hadoop Image Processing Framework is largely a software design project, with a goal of hiding Hadoop's functionality and provide users with a library to effectively use Hadoops MapReduce system for large scale image processing.
-We believe that the ease of use and java oriented semantics will make the process of creating large scale image experiments further more easier.
-This framework is an excellent tool for Hadoop novice users and image processing researchers because it allows development of large scale image processing applications.
+	The Hadoop Image Processing Framework is largely a software engineering platform, with the goal of hiding Hadoop's complexity while providing users with the ability to use the system for large-scale image processing without becoming crack Hadoop engineers.
+The framework's ease of use and Java-oriented semantics will further ease the process of creating large scale image applications and experiments.
+This framework is an excellent tool for novice Hadoop users, image application developers and computer vision researchers, allowing the rapid development of image software that can take advantage of the huge data stores, rich metadata and global reach of current online image sources.
 
 	In the following section  we will describe prior work in this area.
-Next we describe the overview for the Hadoop Image Processing Framework including the Downloader, Processor and Extractor statges.
+Next we present an overview of the Hadoop Image Processing Framework including the Downloader, Processor and Extractor stages.
 Additionally, we describe our approach of distributing tasks for MapReduce.
-Finally we demonstrate the potential of this framework with a few examples of image processing.
+Finally, we demonstrate the potential of this framework with quantitative analysis and experiments performed on image processing tasks.
 
-	\section{Prior Work}
+	\section{Related Work}
 	With rapid increase of usage of online photo storage and social media like facebook, flickr and picasa the amount of image data available is larger than ever before.

### 2015-03-13 Paper.tex
cited references in introduction

@@
 In fact, Hadoop is in use all over the world.
-Studies using Hadoop have been performed, dealing with text data files, analyzing large volumes of DNA sequence data, converting the data of a large number of still images to PDF format, and carrying out feature selection/extraction in astronomy.
+Studies using Hadoop have been performed, dealing with text data files\cite{Lin2010}, analyzing large volumes of DNA sequence data\cite{McKenna2010}, converting the data of a large number of still images to PDF format, and carrying out feature selection/extraction in astronomy\cite{wiley2011}.
 %Cite the papers in which these examples appear
@@
 \section{Conclusion}
-This paper has decribed our framework fo image processing for large scale image processing applications on the MapReduce framework.
+This paper has described our framework fo image processing for large scale image processing applications on the MapReduce framework.
 The framework is designed to abstract the technical details of Hadoop's powerful MapReduce framework and provide a way for users who want to process large image datasets.

### 2015-03-14 Paper.tex
Added a new section image processing computations

@@
 \usepackage{graphicx}
-
+\usepackage{multirow}
 \graphicspath{ {images/} }
@@
 
-\section{Experiments and Observations}
+\section{Image Processing Computations}
+To address the framework capability, we considered to take some existing well used image processing algorihtms and enable running its functionality on a hadoop cluster.
+These considerations led us to a development of image processing functionality in java for the following computations: flat field correction, canny edge detection and image segmentation.
+These computations are computationally intensive and data intensive which are operated on range from hundreds to hundred thousands of images.
+Next we categorize these three computations
 
-We describe two image applications using Hadoop Image Processing framework.
-These applications enable user perform large scale image processing operations with ease.
-These applications are complex and inefficient one existing platforms, yet simple to implement using the framework
+%\subsection{Flat Field Correction}
+%Flat Field Correction(FFC) is described mathematically below:
+%\begin{equation}
+%	I^{FFC}(x,y) = \frac{I^{RAW}(x,y)-DI(x,y)}{WI(x,y)-DI(x,y)}
+%\end{equation}
+
+%where \begin{math}I^{FFC}(x,y))\end{math} is the flat-field corrected image intensity, \begin{math}DI(x,y)\end{math} is the dark image acquired by closing the camera shutter, \begin{math}I^{RAW}(x,y)\end{math} is the raw uncorrected image density and WI(x,y) is the flat field intensity acquired without any object to correct primarily for spatial shading.
+This is the simpleset computations that consists of two subtractions and one division for pixel.
+It needs the DI and WI images co-located from the distributed execution perspective.
+
+\subsection{Laplacian Filter}
+The Laplacian is a 2-D isotropic measure of the 2nd spatial derivative of an image.
+The Laplacian of an image highlights regions of rapid intensity change and is therefore often used for edge detection.
+The Laplacian is often applied to an image that has first been smoothed with something approximating a Gaussian smoothing filter in order to reduce sensitivity to noise.
+The operator normally takes a single gray level image as input and produces another gray level image as output.
+
+The Laplacian \begin{math} L(x,y)\end{math}of an image with pixel intensity values \begin{math}I(x,y)\end{math} is given by 
+\begin{equation}
+L(x,y) = \frac{\partial^2 I}{\partial x^2} + \frac{\partial^2 I}{\partial y^2}
+\end{equation}
+
+This is calculated using a convolution filter.
+
+Since the input image is represented as a set of discrete pixels, we have to find a discrete convolution kernel that can approximate the second derivatives in the definition of the Laplacian.
+This a simplest computation that consists of few additions and multiplications.
 
 \subsection{Canny Edge Detection}
-
 Canny Edge Detection\cite{Canny86} is an multi-stage algorithm to detect a wide range of edges in images.
@@
 
-\begin{figure}[h]
-	\centering
-	\includegraphics[width=0.45\textwidth]{input-canny}
-	\caption{A Sample Image from downloaded image bundle}
-	\label{fig:input-canny}
-\end{figure}
+The process of Canny edge detection algorithm can be broken down to 5 different stages:
 
-\begin{figure}[h]
-	\centering
-	\includegraphics[width=0.45\textwidth]{output-canny}
-	\caption{Processed image using canny edge detection}
-	\label{fig:output-canny}
-\end{figure}
+\begin{itemize}
+	\item Apply Gaussian filter to smooth the image in order to remove the noise.
+	\item Find the intensity gradients of the image.
+	\item Apply non-maximum suppression to get rid of spurious response to edge detection
+	\item Apply double threshold to determine potential edges.
+	\item Track edge by hysteresis. Finalize the detection of edges by suppressing all the other edges that are weak and not connected to strong edges. 
+\end{itemize}
 
-\subsubsection{Characterstics of Image Dataset, Hardware and Software}
-We computed edge detection on a dataset which is about 1TB size.
+\subsection{Image segmentation using k-means clustering}
+K-Means algorithm is an unsupervised clustering algorithm that classifies the input data points into multiple classes based on their inherent distance from each other.
+The algorithm assumes that the data features form a vector space and tries to find natural clustering in them.
+The points are clustered around centroids \begin{math}\mu_{i} \forall i = 1...k\end{math} which are obtained by minimizing the objective
+\begin{equation}
+	V = \sum\limits_{i=1}^n \sum\limits_{x_{j}\in S_{i}} (x_{j} - \mu_{i})^2
+\end{equation}
+
+where there are \begin{math}k\end{math} clusters \begin{math} S_{i},i=1,2,...,k \end{math} and \begin{math} \mu_{i} \end{math} is the centroid or mean point of all the points \begin{math}x_{j} \in S_{i}\end{math}
+
+As a part of the experiment, an iterative version of the algorithm was implemented. The algorithm takes a 2 dimensional image as input. Various steps in the algorithm are as follows:
+
+\begin{itemize}
+	\item Compute the intensity distribution(also called the histogram) of the intensities.
+	\item Initialize the centroids with k random intensities.
+	\item Repeat the following steps until the cluster labels of the image does not change anymore.
+	\item Cluster the points based on distance of their intensities from the centroid intensities.
+	\begin{equation}
+		c^{i} := arg min_{j}||x^{i} - \mu_{j}||^2
+	\end{equation}
+	\item Compute the new centroid for each of the clusters.
+	\begin{equation}
+		\mu_{i} = \frac{\sum\limits_{i=1}^{m} 1\{c_{i} = j\}x^i}{\sum\limits_{i=1}^{m} 1\{c_{i} = j\}}
+	\end{equation}
+
+	where \begin{math}k\end{math} is a parameter of the algorithm (the number of clusters to be found), \begin{math}i\end{math} iterates over that all the intensities, \begin{math}j\end{math} iterates over all the centroids and \begin{math}\mu_{i}\end{math} are the centroid intensities.
+\end{itemize}
+
+\section{Experimental Results}
+
+We describe two image processing algorithms using Hadoop Image Processing framework.
+These image processing algorithms are relatively simple to implement on a single node, implementing these algorithms are complex and inefficient on existing platforms, yet simple to implement using the framework.In the following section we provided image processing computations with the data set described in Section VA.
+The hardware and software specifications are provided in Section VB as well as observations in Section VC
+
+%\begin{figure}[h]
+%	\centering
+%	\includegraphics[width=0.45\textwidth]{input-canny}
+%	\caption{A Sample Image from downloaded image bundle}
+%	\label{fig:input-canny}
+%\end{figure}
+
+%\begin{figure}[h]
+%	\centering
+%	\includegraphics[width=0.45\textwidth]{output-canny}
+%	\caption{Processed image using canny edge detection}
+%	\label{fig:output-canny}
+%\end{figure}
+
+\subsection{Characteristics of Image Dataset, Hardware and Software}
+We computed image processing computations on a dataset which is about 1TB size.
 The data set is made from the flickr search based on a keyword flight.
@@
 The example data set is composed of 220000 images at 4.76 MB/image equals to 1 TB.
-We ran the edge detection on the cluster with 7 nodes having 2 to 4 virtual processors with 4 to 8 GB RAM.
-Each virtual processor clocks at a speed of 3.2 Ghz.
-The algorithm was implemented independently and used on single desktop, Hadoop cluster without framework and Hadoop cluster with framework.
 
-\subsubsection{Observations}
+\subsection{Computer Hardware and Software Characterstics}
+We ran the image processing computations on a cluster with 6 nodes and on a desktop computer.
+Table summarizes the cluster and desktop hardware and software specifications.
+The cluster nodes differ in terms of CPU speed and RAM.
+We installed hadoop, cloudera CDH5 and java 1.7 on the cluster to support java code execution.
+The desktop computer had similar software configuration to the cluster.
+
+\begin{table}[h]
+	\begin{tabular}{|l|l|l|l|}
+		\hline
+		& \multicolumn{1}{c|}{\textbf{Specs}}                            & \multicolumn{1}{c|}{\textbf{Cluster}}                                                                                                           & \multicolumn{1}{c|}{\textbf{Desktop}}                                                                               \\ \hline
+		\multirow{2}{*}{Hardware} & Cluster Nodes                                                  & \begin{tabular}[c]{@{}l@{}}6 datanodes having \\ from 2 to 4 virtual \\ processors with 4 GB\\ RAM and namenode \\ having 8 GB RAM\end{tabular} & \begin{tabular}[c]{@{}l@{}}Intel Xeon @ 2Ghz\\ 8 cores, 16GB of RAM \\ and hyper threading\\ activated\end{tabular} \\ \cline{2-4}
+		& Networking                                                     & 1Gbit/second                                                                                                                                    &                                                                                                                     \\ \hline
+		\multirow{3}{*}{Software} & \begin{tabular}[c]{@{}l@{}}Java Virtual\\ Machine\end{tabular} & \begin{tabular}[c]{@{}l@{}}Java version \\ "1.7.0\_45"\\ Java SE Runtime\\ Environment\end{tabular}                                             & \begin{tabular}[c]{@{}l@{}}Java version \\ "1.7.0\_45"\\ Java SE Runtime\\ Environment\end{tabular}                 \\ \cline{2-4}
+		& Hadoop                                                         & \begin{tabular}[c]{@{}l@{}}hadoop 2.5\\ cloudera CDH5\end{tabular}                                                                              &                                                                                                                     \\ \cline{2-4}
+		& \begin{tabular}[c]{@{}l@{}}Operating\\ System\end{tabular}     & CentOS 6.5                                                                                                                                      & CentOS 6.5                                                                                                          \\ \hline
+	\end{tabular}
+\end{table}
+
+\subsection{Observations}
 
@@
 When the data set size is low, the program runs smooth with best performance.
-As data set size increases to Terabytes the system can't handle such huge amounts of data.
+As data set size increases to Tera-bytes the system can't handle such huge amounts of data.
 To handle such large amounts of data we use the distributed computing using hadoop.

### 2015-03-14 Paper.tex
Further language edits.

@@
 \begin{abstract}
-		With the rapid growth of social media, the number of images being uploaded to the internet is exploding.
-Massive quantities of images are shared through multi-platform services such as Snapchat, Instagram, Facebook and WhatsApp; recent studies estimate that over 1.8 billion photos are uploaded every day.
-However, for the most part, applications that make use of this vast data have yet to emerge.
-Most current image processing applications, designed for small-scale, local computation, do not scale well to web-sized problems with their large requirements for computational resources and storage.
-The emergence of processing frameworks such as the Hadoop MapReduce\cite{Hadoop2009} platform addresses the problem of providing a system for computationally intensive data processing and distributed storage.
-However, to learn the technical complexities of developing useful applications using Hadoop requires a large investment of time and experience on the part of the developer.
-As such, the pool of researchers and programmers with the varied skills to develop applications that can use large sets of images has been limited.
-To address this we have developed the Hadoop Image Processing Framework, which provides a Hadoop-based library to support large-scale image processing.
-The main aim of the framework is to allow developers of image processing applications to leverage the Hadoop MapReduce framework without having to master its technical details and introduce an additional source of complexity and error into their programs.
+With the rapid growth of social media, the number of images being
+uploaded to the internet is exploding. Massive quantities of images
+are shared through multi-platform services such as Snapchat,
+Instagram, Facebook and WhatsApp; recent studies estimate that over
+1.8 billion photos are uploaded every day.  However, for the most
+part, applications that make use of this vast data have yet to
+emerge. Most current image processing applications, designed for
+small-scale, local computation, do not scale well to web-sized
+problems with their large requirements for computational resources and
+storage.  The emergence of processing frameworks such as the Hadoop
+MapReduce\cite{Hadoop2009} platform addresses the problem of providing
+a system for computationally intensive data processing and distributed
+storage. However, to learn the technical complexities of developing
+useful applications using Hadoop requires a large investment of time
+and experience on the part of the developer.  As such, the pool of
+researchers and programmers with the varied skills to develop
+applications that can use large sets of images has been limited. To
+address this we have developed the Hadoop Image Processing Framework,
+which provides a Hadoop-based library to support large-scale image
+processing. The main aim of the framework is to allow developers of
+image processing applications to leverage the Hadoop MapReduce
+framework without having to master its technical details and introduce
+an additional source of complexity and error into their programs.
 \end{abstract}
@@
 \section{Introduction}
-	With the spread of social media in recent years, a large amount of image data has been accumulating.
-When processing this massive data resource has been limited to single computers, computational power and storage ability quickly become bottlenecks.
-Alternately, processing tasks can typically be performed on a distributed system by dividing the task into several subtasks.
-The ability to parallelize tasks allows for scalable, efficient execution of resource-intensive applications.
-The Hadoop MapReduce framework provides a platform for such tasks.
+With the spread of social media in recent years, a large amount of
+image data has been accumulating. When processing this massive data
+resource has been limited to single computers, computational power and
+storage ability quickly become bottlenecks. Alternately, processing
+tasks can typically be performed on a distributed system by dividing
+the task into several subtasks. The ability to parallelize tasks
+allows for scalable, efficient execution of resource-intensive
+applications.  The Hadoop MapReduce framework provides a platform for
+such tasks.
 
-	When considering operations such as face detection, image classification\cite{Li2009} and other types of processing on images, there are limits on what can be done to improve performance of single computers to make them able to process information at the scale of social media.
-Therefore, the advantages of parallel distributed processing of a large image dataset by using the computational resources of a cloud computing environment should be considered.
-In addition, if computational resources can be secured easily and relatively inexpensively, then cloud computing is suitable for handling large image data sets at very low cost and increased performance.
-Hadoop, as a system for processing large numbers of images by parallel and distributed computing, seems promising.
-In fact, Hadoop is in use all over the world.
-Studies using Hadoop have been performed, dealing with text data files\cite{Lin2010}, analyzing large volumes of DNA sequence data\cite{McKenna2010}, converting the data of a large number of still images to PDF format, and carrying out feature selection/extraction in astronomy\cite{wiley2011}.
-%Cite the papers in which these examples appear
-These examples demonstrate the usefulness of the Hadoop system, which can run multiple processes in parallel for load balancing and task management.
+When considering operations such as face detection, image
+classification\cite{Li2009} and other types of processing on images,
+there are limits on what can be done to improve performance of single
+computers to make them able to process information at the scale of
+social media. Therefore, the advantages of parallel distributed
+processing of a large image dataset by using the computational
+resources of a cloud computing environment should be considered. In
+addition, if computational resources can be secured easily and
+relatively inexpensively, then cloud computing is suitable for
+handling large image data sets at very low cost and increased
+performance. Hadoop, as a system for processing large numbers of
+images by parallel and distributed computing, seems promising. In
+fact, Hadoop is in use all over the world. Studies using Hadoop have
+been performed, dealing with text data files\cite{Lin2010}, analyzing
+large volumes of DNA sequence data\cite{McKenna2010}, converting the
+data of a large number of still images to PDF format, and carrying out
+feature selection/extraction in astronomy\cite{wiley2011}.  These
+examples demonstrate the usefulness of the Hadoop system, which can
+run multiple processes in parallel for load balancing and task
+management.
 
-	Most of the image processing applications that use the Hadoop MapReduce framework are highly complex and impose a staggering learning curve.
-The overhead, in programmer time and expertise, required to implement such applications is cumbersome.
-To address this, we present the Hadoop Image Processing Framework, which hides the highly technical details of the Hadoop system and allows programmers who can implement image processing algorithms but who have no particular expertise in distributed systems to nevertheless leverage the advanced resources of a distributed, cloud-oriented Hadoop system.
-Our framework provides users with easy access to large-scale image data, smoothly enabling rapid prototyping and flexible application of complex image processing algorithms to very large, distributed image databases.
+Most of the image processing applications that use the Hadoop
+MapReduce framework are highly complex and impose a staggering
+learning curve.  The overhead, in programmer time and expertise,
+required to implement such applications is cumbersome.  To address
+this, we present the Hadoop Image Processing Framework, which hides
+the highly technical details of the Hadoop system and allows
+programmers who can implement image processing algorithms but who have
+no particular expertise in distributed systems to nevertheless
+leverage the advanced resources of a distributed, cloud-oriented
+Hadoop system. Our framework provides users with easy access to
+large-scale image data, smoothly enabling rapid prototyping and
+flexible application of complex image processing algorithms to very
+large, distributed image databases.
 
-	The Hadoop Image Processing Framework is largely a software engineering platform, with the goal of hiding Hadoop's complexity while providing users with the ability to use the system for large-scale image processing without becoming crack Hadoop engineers.
-The framework's ease of use and Java-oriented semantics will further ease the process of creating large scale image applications and experiments.
-This framework is an excellent tool for novice Hadoop users, image application developers and computer vision researchers, allowing the rapid development of image software that can take advantage of the huge data stores, rich metadata and global reach of current online image sources.
+The Hadoop Image Processing Framework is largely a software
+engineering platform, with the goal of hiding Hadoop's complexity
+while providing users with the ability to use the system for
+large-scale image processing without becoming crack Hadoop engineers.
+The framework's ease of use and Java-oriented semantics will further
+ease the process of creating large scale image applications and
+experiments. This framework is an excellent tool for novice Hadoop
+users, image application developers and computer vision researchers,
+allowing the rapid development of image software that can take
+advantage of the huge data stores, rich metadata and global reach of
+current online image sources.
 
-	In the following section  we will describe prior work in this area.
-Next we present an overview of the Hadoop Image Processing Framework including the Downloader, Processor and Extractor stages.
-Additionally, we describe our approach of distributing tasks for MapReduce.
-Finally, we demonstrate the potential of this framework with quantitative analysis and experiments performed on image processing tasks.
+In the following section we will describe prior work in this
+area. Next we present an overview of the Hadoop Image Processing
+Framework including the Downloader, Processor and Extractor
+stages. Additionally, we describe our approach of distributing tasks
+for MapReduce. Finally, we demonstrate the potential of this framework
+with quantitative analysis and experiments performed on image
+processing tasks.
 
 \section{Related Work}
-	With rapid increase of usage of online photo storage and social media like facebook, flickr and picasa the amount of image data available is larger than ever before.
-The amount of image data is growing rapidly every day.
-According to Buzzfeed for every minute 27800 photos are uploaded to Instagram.
-As per facebook, they get 208,300 photos uploaded every minute\cite{Horaczek2013}.
-This alone provides a plethora of image that can scale up to billions of images.
-It is these reasons that motivate the need for research with image processing applications that take advantage of large sets of images
+With the rapid usage increase of online photo storage and social media
+on sites like Facebook, Flickr and Picasa, more image data is
+available than ever before, and is growing every day.  Every minute
+27800 photos are uploaded to Instagram, while Facebook receives
+208,300 photos over the same time frame.\cite{Horaczek2013} This alone
+provides a source of image data that can scale into the billions.  The
+explosion of available images on social media has motivated image
+processing research and application development that can take
+advantage of very large image data stores.
 
-	A brief overview of previous work sharing the themes of image processing and distributed computation is presented below:
-
-	Web-Scale Computer Vision using MapReduce for Multimedia Data Mining \cite{White2010} present a case study of classifying and clustering billions of regular images using MapReduce.
-A way of pre-processing images for use in a sliding-window approach for object recognition is described.
-Distributed High Performance Video Processing in the cloud, outlines some of the limitations of the MapReduce model when dealing with high-speed video encoding, namely its dependence on the NameNode as a single point of failure, and lack of possibility for generalization in order to suit the issue at hand.
+\cite{White2010} presents a case study of classifying and clustering billions of regular images using MapReduce.
+It describes an image pre-processing technique for use in a sliding-window approach for object recognition.
+Distributed High Performance Video Processing in the cloud outlines some of the limitations of the MapReduce model when dealing with high-speed video encoding, namely its dependence on the NameNode as a single point of failure, and lack of possibility for generalization in order to suit the issue at hand.
 An alternate optimized implementation is proposed for providing cloud-based IaaS (Infrastructure as a Service) Solution.
@@
 \section{Methodology}
-Hadoop Image processing Framework is created to provide users a tool to develop a large scale image processing applications with ease.
+The Hadoop Image Processing Framework is intended to provide users
+with an accessible, easy-to-use tool for developing large-scale image
+processing applications.
 
-The main goals of Hadoop Image Processing framework:
+The main goals of the Hadoop Image Processing Framework are:
 \begin{itemize}
-	\item Provide an open source framework over Hadoop MapReduce for developing large scale image applications 
-	\item Flexibility to store images in various Hadoop File Formats
-	\item Present users with an intuitive interface for image based operations which are highly parallelized and balanced and hide the high technical details of hadoop MapReduce Framework
-	\item Allow interoperability between various image processing libraries
+\item Provide an open source framework over Hadoop MapReduce for
+  developing large-scale image applications
+\item Provide the ability to flexibly store images in various Hadoop
+  file formats
+\item Present users with an intuitive application programming
+  interface for image-based operations which is highly parallelized
+  and balanced, but which hides the technical details of Hadoop
+  MapReduce
+\item Allow interoperability between various image processing
+  libraries
 \end{itemize}
@@
 \subsection{Downloading and storing image data}
-Hadoop's distributed file system(HDFS) to store files in various nodes throughout the cluster.
-Storing small files\cite{White2009} is a big problem in Hadoop.
-A small file is one which is significantly smaller than HDFS block size.
-As the large image dataset can have many number of small files and the problem is HDFS cant handle lots of files.
-The problem can be solved by providing some kind of container to group the files in some way.
-Hadoop offers a few options:
+Hadoop uses the Hadoop Distributed File System (HDFS) to store files
+in various nodes throughout the cluster.  One of Hadoop's significant
+problems is that of small file storage. \cite{White2009} A small file
+is one which is significantly smaller than HDFS block size.  Large
+image datasets are made up of small image files in great numbers,
+which is a situation which HDFS has a great deal of trouble
+handling. The problem can be solved by providing a container to group
+the files in some way. Hadoop offers a few options:
 \begin{itemize}
@@
 
-The Downloader Module of the framework does the following work
+The Downloader Module of our Hadoop Image Processing Framework
+performs the following operations:
 

### 2015-03-15 Paper.tex
More copyedits.

@@
 
-\textbf{Step 1: Input a URL List.} Initially user inputs a file containing url's of images to download.
-The input list should be a text file with one image url per line.
-The list can be generated by hand, database or a search.
-The framework provides a extendable ripper module to extract urls from flickr, google and databases.
-In addition to the list the user will input the type of image bundle to be generated(e.g.
-har, sequence or map).
-According to the input, the urls are divided across the nodes for maximum efficiency and parallelism to download the images.
-Each node map task will generate few image bundles as per the input type containing all the of the images it downloaded.
-In the reduce phase, the Reducer will merge these image bundles  into a large image bundle.
+\textbf{Step 1: Input a URL List.} Initially users input a file
+containing URLs of images to download. The input list should be a text
+file with one image URL per line. The list can be generated by hand,
+extracted from a database or a provided by a search. The framework
+provides a extendable ripper module for extracting URLS from Flickr
+and Google image searches and from SQL databases. In addition to the
+list the user selects the type of image bundle to be generated
+(e.g. HAR, sequence or map).  Our system divides the URLs for download
+across the available processing nodes for maximum efficiency and
+parallelism.  Each node map task generates several image bundles
+appropriate to the selected input type, containing all of the images
+it downloaded. In the reduce phase, the Reducer will merge these image
+bundles into a large image bundle.
 
-\textbf{Step 2: Split URLs across nodes.}
-From the input, file containing list of image urls and the type of file to be generated, we equally distribute the task of downloading images across the all the nodes in the cluster.
-The nodes are efficiently managed so that no memory overflow can occur even for terabytes of image downloaded by a single map task.
-This allows maximum parallelization for downloading.
-Image urls are distributed to all nodes equally and different map tasks begin downloading the image set it is responsible for.
+\textbf{Step 2: Split URLs across nodes.}  From the input file
+containing the list of image URLs and the type of file to be
+generated, we equally distribute the task of downloading images across
+the all the nodes in the cluster. The nodes are efficiently managed so
+that no memory overflow can occur even for terabytes of images
+downloaded in a single map task. This allows maximum downloading
+parallelization. Image URLs are distributed among all available
+processing nodes, and each map task begins downloading its respective
+image set.
 
@@
 	\includegraphics[width=0.45\textwidth]{down-map}
-	\caption{Individual Map task of Downloader Module}
+	\caption{Individual map task of Downloader Module}
 	\label{fig:down-map}
@@
 
-\textbf{Step 3: Download image data from URLs.}
-For every url retrieved in the map task, a connection is established as per transfer protocal (e.g.
-ftp, http, https etc.).
-Once connected, we check the file type and if its a valid image we get an InputStream to the connection.
-From this InputStream, we generate a new HImage object and add the image to the image bundle.
-The HImage class holds the image data and helps user to play with image and image header data.
-HImage class also provides an interoperability between various image data types (e.g.
-BufferedImage, Mat etc.,).
+\textbf{Step 3: Download image data from URLs.}  For every URL
+retrieved in the map task, a connection is established according to
+the appropriate transfer protocol (e.g. FTP, HTTP, HTTPS, etc.). Once
+connected, we check the file type.  Valid images are assigned
+InputStreams associated with the connection. From these InputStreams,
+we generate new HImage objects and add the images to the image
+bundle. The HImage class holds the image data and provides an
+interface for the user's manipulation of image and image header
+data. The HImage class also provides interoperability between various
+image data types (e.g. BufferedImage, Mat, etc.).
 
-\textbf{ Step 4: Store images in an image bundle. } 
-Once HImage object is received, we can add the image to the image bundle simply by passing the HImage object to the appendImage method.
-Each map task generates a few or many image bundles depending on the image list.
-In the reduce phase, all these image bundles are merged into one large bundle.
+\textbf{ Step 4: Store images in an image bundle. }  Once an HImage
+object is received, it can be added to the image bundle simply by
+passing the HImage object to the appendImage method. Each map task
+generates a number of image bundles depending on the image list. In
+the reduce phase, all of these image bundles are merged into one large
+bundle.
 
@@
 	\includegraphics[width=0.45\textwidth]{down-node}
-	\caption{Single Node running the Downloader Module handled by the framework and transparent to user}
+	\caption{Single node running the Downloader Module (handled by
+          the framework and transparent to the user)}
 	\label{fig:down-node}
@@
 \subsection{Processing image bundle using MapReduce}
-Hadoop MapReduce programs handle input and output data very efficiently, but the main problem is representing the image data.
-For instance, to distribute images across map nodes would require to pass images as string, then decode image into a specified format to access pixel information which is not only inefficient but also inconvenient.
-To overcome this problem we want to able to represent the images as many different formats to increase flexibility.
-The framework focuses on bringing familiar data types directly to user.
+Hadoop MapReduce programs handle input and output data very
+efficiently, but their native data exchange formats are not convenient
+for representing or manipulating image data.  For instance,
+distributing images across map nodes require the translation of images
+into strings, then later decoding these image strings into specified
+formats in order to access pixel information.  This is both
+inefficient and inconvenient. To overcome this problem, images should
+be represented in as many different formats as possible, increasing
+flexibility. The framework focuses on bringing familiar data types
+directly to user.
 
-As distribution is important in MapReduce, we want to process the image in the same machine where the bundle block resides.
-In general, the user needs to create an InputFormat and Record Reader classes to specify the MapReduce job and distribute the input among nodes.
-This task is a bit cumbersome.
-In order to make the framework more easier we took care of the InputFormat and RecordReaders.
+As distribution is important in MapReduce, images should be processed
+in the same machine where the bundle block resides. In a generic
+MapReduce system, the user is responsible for creating InputFormat and
+RecordReader classes to specify the MapReduce job and distribute the
+input among nodes. This is a cumbersome and error-prone task; the
+Hadoop Image Processing Framework provides such InputFormat and
+RecordReaders for the sake of the system's ease of use.
 
-Images are distributed as various image data types and users can immediately have access to pixel values.
-As we attempt to obtain the pixel values we lose the valuable image header data(e.g.
-jpeg exif data, IHDR\cite{David03} Image Header).
-The framework, holds the image data in a special HImageHeader data type before converting the image bytes into pixel values.
-After working on the pixel data, it was made sure that image header is appended to the processed pixels.
-Even though it adds an storage overhead, image header data is preserved even after processing.
+Images are distributed as various image data types and users have
+immediate access to pixel values.  If pixel values are naively
+extracted from the image byte formats, valuable image header data
+(e.g. JPEG EXIF data or IHDR \cite{David03} image headers) is
+lost. The framework holds the image data in a special HImageHeader
+data type before converting the image bytes into pixel values.  After
+processing pixel data, image headers are reattached to the processed
+results.  The small amount of storage overhead required for this
+functionality is a trade-off worth making for the sake of preserving
+image header data after processing.
 
-The working of Processor module of the framework is described below:
+The functionality of the framework's Processor module is described
+below:
 
-\textbf{Step 1: Input the Algorithm.} We assume that the user writes an algorithm which extends a Generic Algorithm class.
-This class is passed as an argument to the processor module.
-The framework starts a MapReduce Job with the algorithm as an input.
-The GenericAlgorithm holds an HImage variable, this helps user to write an algorithm on the single image which iterates over the whole image bundle.
-In addition to the algorithm the user should input the image bundle file that needs to be processed.
-According to the input the image bundle is divided across nodes as individual map task.
-Each Map task will process the image as of the algorithm and appends them to the output image bundle.
-In the reduce phase, the Reducer will merge these image bundles into a large image bundle.
+\textbf{Step 1: Devise the algorithm.} We assume that the user writes
+an algorithm which extends the provided GenericAlgorithm class. This
+class is passed as an argument to the processor module. The framework
+starts a MapReduce job with the algorithm as an input. The
+GenericAlgorithm holds an HImage variable; this allows user to write
+an algorithm on a single image data structure, which the framework
+then iterates over the entire image bundle. In addition to the
+algorithm, the user should provide the image bundle file that needs to
+be processed.  Depending on the specifics of the image bundle
+organization and contents, the bundle is divided across nodes as
+individual map tasks. Each map task will apply the processing
+algorithm to each local image and append them in turnto the output
+image bundle. In the reduce phase, the Reducer merges these image
+bundles into a large image bundle.
 
-\textbf{Step 2: Split Imagebundle across nodes.} The input image bundle is stored in HDFS as blocks.
-In order to obtain maximum throughput it was made sure the map task is run in the same the block where it resides, using custom input format and record reader classes.
-This allows maximum parallelization without a problem of transferring data across nodes.
-The image bundle now executes different map tasks for the image data it is responsible for.
+\textbf{Step 2: Split image bundle across nodes.} The input image
+bundle is stored as blocks in the HDFS.  In order to obtain maximum
+throughput, the framework establishes each map task to run in the same
+block where it resides, using custom input format and record reader
+classes. This allows maximum parallelization without the problem of
+transferring data across nodes.  Each image bundle now applies
+different map tasks to the image data for which it is responsible.
 
@@
 	\includegraphics[width=0.45\textwidth]{pro-map}
-	\caption{Individual Map task of Processor}
+	\caption{Individual map task of Processor Module}
 	\label{fig:pro-map}

### 2015-03-15 Paper.tex
Finished editing methodology section.

@@
 
-\textbf{Step 3: Process individual image.} For every HImage retrieved in the map task, the algorithm is run on the HImage which is given as the input to the processor module.
-The HImage retrieves the image as data type(e.g.
-BufferedImage, Mat image) in the algorithm.
-Once the image data type is retrieved processing takes place.
-After processing, the image header data of the original image is appended to the processed image, thereby preserving image header data.
-The processed image is appended to the temporary bundle generated by the map task.
+\textbf{Step 3: Process individual image.}  The processing algorithm
+devised by the user and provided as input to the Processing Module is
+applied to every HImage retrieved in the map task.  The HImage
+provides its image data in the data format (e.g. Java BufferedImage,
+OpenCV Mat, etc.) requested by the user and used by the processing
+algorithm. Once the image data type is retrieved, processing takes
+place. After processing, the preserved image header data from the
+original image is appended to the processed image. The processed image
+is appended to the temporary bundle generated by the map task.
 
-\textbf{Step 4: Store processed images in an image Bundle.} Every map task generate a image bundle after the map task is completed.
-Once the map phase is completed there are many bundles.
-In the reduce phase, all these temporary image bundles are merged into a single large file which contains all the processed images.
+\textbf{Step 4: Store processed images in an image bundle.} Every map
+task generates an image bundle upon completion of its processing.
+Once the map phase is completed there are many bundles scattered
+across the computing cluster.  In the reduce phase, all of these
+temporary image bundles are merged into a single large file which
+contains all the processed images.
 
@@
   \includegraphics[width=0.45\textwidth]{pro-node}
-	\caption{Single Node running the Processor Module handled by the framework and transparent to user}
+  \caption{Single node running the Processor Module (handled by
+    the framework and transparent to user)}
   \label{fig:pro-node}
@@
 
-\subsection{Extracting image bundle using MapReduce}
-The framework creates and processes image bundles but there should be a way to extract these images and make them viewable.
-Generally, Hadoop extracts images from the image bundle iteratively using a single node which is not efficient.
-To make the process distributed among different nodes we designed an extractor module which can extract images in parallel in all the nodes.
+\subsection{Extracting image bundles using MapReduce}
+In addition to creating and processing image bundles, the framework
+provides a method for extracting and viewing these images.  Generally,
+Hadoop extracts images from an image bundle iteratively, inefficiently
+using a single node for the task. To address this inefficiency, we
+designed an Extractor module which extracts images in parallel across
+all available nodes.
 
-Distribution plays a key role in MapReduce, we want to make use of the nodes effectively to make the process of extraction efficient.
-In general, the user needs to create custom InputFormat and RecordReader classes to specify the MapReduce job and distribute the input among nodes.
-In order to over come this the framework takes care of these Custom InputFormat and RecordReader to make working with hadoop more easier.
+Distribution plays a key role in MapReduce; we want to make effective
+and efficient use of the nodes in the computing cluster.  As
+previously mentioned in the description of the Processor module, a
+user working in a generic Hadoop system must again devise custom
+InputFormat and RecordReader classes in order to facilitate
+distributed image extraction.  The Hadoop Image Processing Framework
+provides this functionality for the extraction task as well, providing
+much greater ease of use for the development of image processing
+applications.
 
-There is often a confusion to where the images need to be extracted.
-The images can be extracted to the local file system of the node or hadoop file system.
-The framework overcomes the problem by providing a flexibility to choose.
+Organizing and specifying the final location of extracted images in a
+large distributed task can be confusing and difficult.  Our framework
+provides this functionality, and allows the user simply to specify
+whether images should be extracted to a local file system or reside on
+the Hadoop DFS.
 
-The working of Extractor module is explained below.
+The process of the Extractor module is explained below:
 
-\textbf{Step 1: Input the image bundle to be extracted.} The image bundle that needs to extracted is passed as a parameter to the Extractor module.
-Along with it the user can send an optional parameter to where the images need to be extracted (defaults to local file system).
-After the input the image bundle is distributed across nodes as individual map task.
-Each map task will extract the image in to the filesystem.
+\textbf{Step 1: Input the image bundle to be extracted.} The image
+bundle specified for extraction is passed as a parameter to the
+Extractor module.  In addition, the user can include an optional
+parameter specifying the images' final location (defaults to local
+file system).  The image bundle is then distributed across the nodes
+as individual map tasks. Each map task will extract the requisite
+images onto the specified filesystem.
 
@@
   \includegraphics[width=0.45\textwidth]{ext-map}
-	\caption{Individual Map Task of Extractor Module}
+  \caption{Individual map task of Extractor Module}
   \label{fig:ext-map}
@@
 
-\textbf{Step 2: Split Image bundle across nodes}
-The input image bundle split across nodes using the custom input format and record reader classes for maximum throughput. Once a Map task starts a HImage object is retrieved.
+\textbf{Step 2: Split Image bundle across nodes} The input image
+bundle is split across available nodes using the framework's custom
+input format and record reader classes for maximum throughput. Once a
+map task starts, HImage objects are retrieved.
 
-\textbf{Step 3: Extract individual image}
-The obtained HImage object is retrieved as image bytes and the image bytes are stored into the file based on its image type. The Extractor module doesn't have a reduce phase.
+\textbf{Step 3: Extract individual image} The image bytes obtained
+from each HImage object are stored onto the filesystem in the
+appropriate file format based on its image type.  The Extractor module
+lacks a reduce phase.
 
@@
   \includegraphics[width=0.45\textwidth]{ext-node}
-	\caption{Single node running the Extractor Module handled by the framework and transparent to user}
+  \caption{Single node running the Extractor Module (handled by the
+    framework and transparent to the user)}
   \label{fig:ext-node}

### 2015-03-15 Paper.tex
Updated charts and observations section

@@
 
-\subsection{Observations}
-
 The desktop implementation of the Algorithm is a java program for processing a small image dataset.
@@
 To handle such large amounts of data we use the distributed computing using hadoop.
-
-\begin{figure}[h]
-	\centering
-	\includegraphics[width=0.45\textwidth]{module-chart}
-	\caption{Module Chart}
-	\label{fig:module-chart}
-\end{figure}
 
@@
 
+\subsection{Observations}
+
+We have documented the following samples in the space of image processing
+computations, and hardware and software configurations applied to the selected
+data set.
+\begin{itemize}
+	\item Three image processing computations described in Table 1.
+	\item Sampled data set of different image counts.
+	\item Hardware cluster configurations used for computations. 
+	\item Coding complexities in different environments for the three image processing computations
+\end{itemize}
+
+The numerical results are shown in Figure 1 through 6.
+
+Figure \ref{fig:perf-chart} illustrates the performance comparison of Single Node, Hadoop using framework and Hadoop without using the framework for Canny Edge Detection.
+
+\begin{figure}[h]
+	\centering
+	\includegraphics[width=0.45\textwidth]{perf-graph}
+	\caption{Canny Edge Detection computations executed on single node and on a Hadoop cluster with and without using framework}
+	\label{fig:perf-chart}
+\end{figure}
+It can be observed that the performance hadoop cluster is almost the same with and without using framework.
+The framework performs well sometimes because of multi-threading is implemented few tasks.
+Considering the single node performance it can be seen that the single node outperforms hadoop if the input is small as the input size increases handling huge data becomes infeasible.
+The same pattern is observed with any image processing algorithm.
+
+Figure \ref{fig:module-chart} compares coding complexities of different modules the framework described.
+The Downloader module and Extractor module are simple to use and require less number of lines of code to configure.
+The processor module includes the algorithm the user devised, which adds up a few extra lines of code.
+but configuring the devised algorithm to a MapReduce job is fairly simple with very less number of steps.
+The modules are included with different methods for flexibility.
+
+\begin{figure}[h]
+	\centering
+	\includegraphics[width=0.45\textwidth]{module-chart2}
+	\caption{Comparing coding complexities of modules in different environments}
+	\label{fig:module-chart}
+\end{figure}
+
+Figure \ref{fig:comp-chart} shows the coding complexities of three image processing computations from Table 1.
+Single node and the framework show similar coding complexity outperforming the image processing without framework.
+The reason behind this is Hadoop Image Processing Framework hides the inner details of MapReduce programming which makes an user to think that he is working on the single node but actually the task is processed in hadoop.
+Without the framework user needs to write a MapReduce job which is cumbersome.
+
+\begin{figure}[h]
+	\centering
+	\includegraphics[width=0.45\textwidth]{comp-chart}
+	\caption{Comparing coding complexities of laplacian filter, canny edge detection and k-means clustering  in different environments}
+	\label{fig:comp-chart}
+\end{figure}
+
 Hadoop Image processing framework provides transparency over two levels.
@@
 Making framework similar to java helps understanding the framework more easier and reduces complexity.
+Listing 2 shows a sample code for setting image headers for processed images.
 
@@
 	\centering
-	\includegraphics[width=0.45\textwidth]{files-chart}
+	\includegraphics[width=0.45\textwidth]{files-chart2}
 	\caption{Comparing File Writers and readers without and with using framework }
@@
 
+\begin{lstlisting}[caption = Setting image headers for processed images using Hadoop Image Processing Framework]
+
+//input - HImage is sent as input
+
+CannyEdge canny = new CannyEdge(input);
+canny.process();
+HImage himage = canny.getProcessedImage();
+himage.setImageHeader(input.getImageHeader());
+
+\end{lstlisting}
 \section{Conclusion}

### 2015-03-15 Paper.tex
Finished editing algorithm descriptions.

@@
 \section{Image Processing Computations}
-To address the framework capability, we considered to take some existing well used image processing algorihtms and enable running its functionality on a hadoop cluster.
-These considerations led us to a development of image processing functionality in java for the following computations: flat field correction, canny edge detection and image segmentation.
-These computations are computationally intensive and data intensive which are operated on range from hundreds to hundred thousands of images.
-Next we categorize these three computations
+To explore the Hadoop Image Processing Framework's capabilities and
+performance, we implemented multiple variations of existing
+widely-used image processing algorithms:
+\begin{itemize}
+\item A stand-alone implementation running on a single node with no
+  distributed processing capability
+\item A generic Hadoop implementation which requires a high level of
+  software engineering expertise within the Hadoop framework
+\item An implementation that employs the Hadoop Image Processing
+  Framework
+\end{itemize}
+
+We chose the following algorithms: Laplacian filtering, Canny edge
+detection and $k$-means image segmentation.  These are widely-used,
+computation- and data-intensive algorithms of varying complexity which
+require large distributed systems for timely operation on hundreds of
+thousands of images.
 
@@
 \subsection{Laplacian Filter}
-The Laplacian is a 2-D isotropic measure of the 2nd spatial derivative of an image.
-The Laplacian of an image highlights regions of rapid intensity change and is therefore often used for edge detection.
-The Laplacian is often applied to an image that has first been smoothed with something approximating a Gaussian smoothing filter in order to reduce sensitivity to noise.
-The operator normally takes a single gray level image as input and produces another gray level image as output.
+The Laplacian is a 2-D isotropic measure of the 2nd spatial derivative
+of an image. The Laplacian of an image highlights regions of rapid
+intensity change and is therefore often used for edge detection. The
+Laplacian is often applied to an image that has first been smoothed
+with an approximation of a Gaussian filter in order to reduce
+sensitivity to noise. The operator normally takes a single gray level
+image as input and produces another gray level image as output.
 
-The Laplacian \begin{math} L(x,y)\end{math}of an image with pixel intensity values \begin{math}I(x,y)\end{math} is given by 
+The Laplacian \begin{math} L(x,y)\end{math}of an image with pixel
+intensity values \begin{math}I(x,y)\end{math} is given by
 \begin{equation}
@@
 
-Since the input image is represented as a set of discrete pixels, we have to find a discrete convolution kernel that can approximate the second derivatives in the definition of the Laplacian.
-This a simplest computation that consists of few additions and multiplications.
+Since the input image is represented as a set of discrete pixels, we
+have to find a discrete convolution kernel that can approximate the
+second derivatives in the definition of the Laplacian. This is a very
+simple computation that consists of a few additions and
+multiplications.
 
 \subsection{Canny Edge Detection}
-Canny Edge Detection\cite{Canny86} is an multi-stage algorithm to detect a wide range of edges in images.
-Edge detection, especially step edge detection is an important technique to extract useful structural information from different vision objects and dramatically reduce the amount of data to be processed.
-Among the edge detection methods developed so far, canny edge detection algorithm is one of the most strictly defined methods that provides good and reliable detection.
+Canny edge detection\cite{Canny86} is an multi-stage algorithm to
+detect a wide range of edges in images. Edge detection, especially
+step edge detection, is an important technique to extract useful
+structural information from visual information, dramatically reducing
+the amount of data to be processed. Among existing edge detection
+methods, the Canny edge detection algorithm is considered to provide
+reliable edge detection without relying on context-specific
+heuristics.
 
-The process of Canny edge detection algorithm can be broken down to 5 different stages:
+The process of the Canny edge detection algorithm can defined in five
+stages:
 
 \begin{itemize}
-	\item Apply Gaussian filter to smooth the image in order to remove the noise.
+\item Apply a Gaussian filter to smooth the image and remove noise.
 \item Find the intensity gradients of the image.
-	\item Apply non-maximum suppression to get rid of spurious response to edge detection
-	\item Apply double threshold to determine potential edges.
-	\item Track edge by hysteresis. Finalize the detection of edges by suppressing all the other edges that are weak and not connected to strong edges. 
+\item Apply non-maximum suppression to get rid of spurious
+  responses to edge detection
+\item Apply a double threshold to determine potential edges.
+\item Track edges by hysteresis. Finalize the detection of
+  edges by suppressing all edges that are ``weak'' and not
+  connected to strong edges.
 \end{itemize}
 
-\subsection{Image segmentation using k-means clustering}
-K-Means algorithm is an unsupervised clustering algorithm that classifies the input data points into multiple classes based on their inherent distance from each other.
-The algorithm assumes that the data features form a vector space and tries to find natural clustering in them.
-The points are clustered around centroids \begin{math}\mu_{i} \forall i = 1...k\end{math} which are obtained by minimizing the objective
+\subsection{Image segmentation using $k$-means clustering}
+The $k$-means algorithm is an unsupervised clustering algorithm that
+classifies input data points into multiple classes based on their
+Euclidean distance from each other. The algorithm assumes that the
+data features form a vector space and tries to find natural
+clusterings within it. The points are clustered around
+centroids \begin{math}\mu_{i} \forall i = 1...k\end{math} which are
+  obtained by minimizing the objective
 \begin{equation}
@@
 
-where there are \begin{math}k\end{math} clusters \begin{math} S_{i},i=1,2,...,k \end{math} and \begin{math} \mu_{i} \end{math} is the centroid or mean point of all the points \begin{math}x_{j} \in S_{i}\end{math}
+where there are \begin{math}k\end{math} clusters \begin{math}
+    S_{i},i=1,2,...,k \end{math} and \begin{math} \mu_{i} \end{math}
+  is the centroid or mean point of all the points \begin{math}x_{j}
+    \in S_{i}\end{math}.
 
-As a part of the experiment, an iterative version of the algorithm was implemented. The algorithm takes a 2 dimensional image as input. Various steps in the algorithm are as follows:
+For the purposes of our experiment, we implemented iterative versions
+of the algorithm. The algorithm takes a 2-dimensional image as
+input. The steps of the algorithm are as follows:
 
 \begin{itemize}
-	\item Compute the intensity distribution(also called the histogram) of the intensities.
-	\item Initialize the centroids with k random intensities.
-	\item Repeat the following steps until the cluster labels of the image does not change anymore.
-	\item Cluster the points based on distance of their intensities from the centroid intensities.
+\item Compute the intensity distribution (also called the
+  histogram) of the image.
+\item Initialize the centroids with $k$ random intensities.
+\item Repeat the following steps until the cluster labels of
+  the image stop changing from one iteration to the next:
+  \begin{itemize}
+  \item Cluster the points based on the distance of their intensities
+    from the centroid's intensity.
 	\begin{equation}
-		c^{i} := arg min_{j}||x^{i} - \mu_{j}||^2
+		c^{i} := \mbox{argmin}_{j}||x^{i} - \mu_{j}||^2
 	\end{equation}
@@
 
-	where \begin{math}k\end{math} is a parameter of the algorithm (the number of clusters to be found), \begin{math}i\end{math} iterates over that all the intensities, \begin{math}j\end{math} iterates over all the centroids and \begin{math}\mu_{i}\end{math} are the centroid intensities.
+	where \begin{math}k\end{math} is a parameter of the algorithm
+          (the number of clusters to be
+          found), \begin{math}i\end{math} iterates over all the
+            intensities, \begin{math}j\end{math} iterates over all the
+              centroids and \begin{math}\mu_{i}\end{math} are the
+              centroid intensities.
+  \end{itemize}
 \end{itemize}

### 2015-03-15 Paper.tex
Updated related work and references

@@
 available than ever before, and is growing every day.  Every minute
-27800 photos are uploaded to Instagram, while Facebook receives
-208,300 photos over the same time frame.\cite{Horaczek2013} This alone
+27800 photos are uploaded to Instagram, \cite{Horaczek2013}while Facebook receives
+208,300 photos over the same time frame. This alone
 provides a source of image data that can scale into the billions.  The
@@
 
-\cite{White2010} presents a case study of classifying and clustering billions of regular images using MapReduce.
+White et.al \cite{White2010} presents a case study of classifying and clustering billions of regular images using MapReduce.
 It describes an image pre-processing technique for use in a sliding-window approach for object recognition.
-Distributed High Performance Video Processing in the cloud outlines some of the limitations of the MapReduce model when dealing with high-speed video encoding, namely its dependence on the NameNode as a single point of failure, and lack of possibility for generalization in order to suit the issue at hand.
-An alternate optimized implementation is proposed for providing cloud-based IaaS (Infrastructure as a Service) Solution.
-In Parallel K-Means clustering of Remote Sensing Images Based on MapReduce \cite{Lv2010} describe using the k-means algorithm in conjunction with MapReduce and satellite/aerophoto images in order to find different elements based on their color.
-Using Transaction Based Parallel computing to solve image processing and computational physics problems by Harold describes the use of distributed computing with two examples video processing and subsurface transport.
-There was no information provided on how image processing parts of the examples.
+Pereira et.al \cite{Pereira2010} outlines some of the limitations of the MapReduce model when dealing with high-speed video encoding, namely its dependence on the NameNode as a single point of failure, and lack of possibility for generalization in order to suit the issue at hand.
+An alternate optimized implementation is proposed for providing cloud-based IaaS (Infrastructure as a Service) solution.
+Zhenhua Lv et.al \cite{Lv2010} describe using the k-means algorithm in conjunction with MapReduce and satellite/aerophoto images in order to find different elements based on their color.
 
-	Case Study of Scientific Data Processing on a Cloud Using Hadoop \cite{Zhang2010} present methods used for processing sequences of microscope images of live cells.
-The images are relatively small 512x512 16-bit pixels, stored in 90 MB folders there were some issues regarding to fitting into Hadoop DFS blocks with were solved by custom InputFormat, InputSplit and Record Reader classes.
-A scalable image processing framework for gigapixel Mars and other celestial body images by Mark Powell describes the way NASA handles image processing of celestial images captured by the Mars Orbiter and rovers.
+Zhang et.al \cite{Zhang2010} present methods used for processing sequences of microscope images of live cells.
+The images are relatively small 512x512 16-bit pixels, stored in 90 MB folders there were some issues regarding to fitting into Hadoop DFS blocks with were solved by custom InputFormat, InputSplit and RecordReader classes.
+Mark Powell et.al \cite{Powell2010} describes the way NASA handles image processing of celestial images captured by the Mars Orbiter and rovers.
 Clear and concise descriptions are provided about the segmentation of gigapixel images into tiles, how the tiles are processed and how image processing framework handles scaling and works with the distributed processing.
-Ultra-fast processing of gigapixel tissue microarray images using high performance computing \cite{Wang2011} about the speeding up the analysis of Tissue MicroArray images by substituting human expert analysis for automated processing algorithms.
+Wang, Yinhai and McCleary\cite{Wang2011} about the speeding up the analysis of Tissue MicroArray images by substituting human expert analysis for automated processing algorithms.
 While the images were measured to be in gigapixels, the content is easily segmented and there was no need to focus on being able to analyze all of image at once.
@@
 
-	Terabyte size image computations on Hadoop cluster \cite{Bajcsy2013} platforms present a characterization of four basic terabyte size image computations on a Hadoop Cluster in terms of their relative efficiency according to the modified Amdahls law.
+Bajcsy et.al \cite{Bajcsy2013} present a characterization of four basic terabyte size image computations on a Hadoop Cluster in terms of their relative efficiency according to the modified Amdahls law.
 The work is motivated by the fact that there is a lack of 6 standard benchmarks and stress tests for big image processing operations on Hadoop framework.
-Terabyte-scale image search\cite{Moise2013} outlines about querying thousands of images in one run using Hadoop MapReduce framework using the MapReduced based eCP Algorithm.
+Moise et.al \cite{Moise2013} outlines about querying thousands of images in one run using Hadoop MapReduce framework using the MapReduced based eCP Algorithm.
 The experiment does an image search on 110 million images collected from web on Grid 5000 Platform.

### 2015-03-15 Paper.tex
Nearly finished with all editing except for related work.

--- 2015-03-15 Paper.tex
@@
 \section{Methodology}
+\label{methodology}
 The Hadoop Image Processing Framework is intended to provide users
@@
 
-\section{Image Processing Computations}
+\section{Image Processing Algorithms}
+\label{algorithms}
 To explore the Hadoop Image Processing Framework's capabilities and
@@
 
-We describe two image processing algorithms using Hadoop Image Processing framework.
-These image processing algorithms are relatively simple to implement on a single node, implementing these algorithms are complex and inefficient on existing platforms, yet simple to implement using the framework.In the following section we provided image processing computations with the data set described in Section VA.
-The hardware and software specifications are provided in Section VB as well as observations in Section VC
+We present execution time and programming complexity results for the
+three image processing algorithms described above, demonstrating the
+effectiveness of the Hadoop Image Processing Framework.  These
+algorithms are not particularly challenging to implement on a single
+node, but producing a distributed Hadoop implementation is quite
+challenging and requires a great deal of familiarity with Hadoop
+concepts and distributed software engineering expertise.  On the other
+hand, our framework provides all of the performance and scalability
+benefits of the Hadoop system while retaining the simplicity of the
+single-node implementation.  In this section we describe the dataset
+and hardware and software specifications used in our experiments, and
+discuss the results.
 
@@
 \subsection{Characteristics of Image Dataset, Hardware and Software}
-We computed image processing computations on a dataset which is about 1TB size.
-The data set is made from the flickr search based on a keyword flight.
-The data set is extracted using the a FlickrRipper in the framework which implements Ripper Interface.
-The Ripper interface can be implemented for any custom Rippers to get the image URLs.
-The example data set is composed of 220000 images at 4.76 MB/image equals to 1 TB.
+We performed image processing computations on a dataset of about 1TB
+in size. The dataset was obtained by performing a Flickr search based
+on the keyword ``flight''.  The dataset was extracted using the
+FlickrRipper, an instance of the Ripper interface provided by our
+Hadoop Image Processing Framework.  The framework currently provides
+rippers for Flickr and Google Image searches, and more will be
+provided in the future.  A Ripper interface extracts image URLs from
+search results, and can be implemented for any set of results in any
+format.  The experimental dataset is composed of 220000 images at 4.76
+MB/image, or about 1 TB.
 
 \subsection{Computer Hardware and Software Characterstics}
-We ran the image processing computations on a cluster with 6 nodes and on a desktop computer.
-Table summarizes the cluster and desktop hardware and software specifications.
-The cluster nodes differ in terms of CPU speed and RAM.
-We installed hadoop, cloudera CDH5 and java 1.7 on the cluster to support java code execution.
-The desktop computer had similar software configuration to the cluster.
+We ran the image processing computations on a 6-node distributed
+cluster and on a single-node desktop computer. The table below
+summarizes the cluster and desktop hardware and software
+specifications. The cluster nodes differ in terms of CPU speed and
+RAM. We installed Hadoop, Cloudera CDH5 and Java 1.7 on the cluster to
+support Java code execution. The desktop computer had a similar
+software configuration to the cluster.
 
@@
 		& \multicolumn{1}{c|}{\textbf{Specs}}                            & \multicolumn{1}{c|}{\textbf{Cluster}}                                                                                                           & \multicolumn{1}{c|}{\textbf{Desktop}}                                                                               \\ \hline
-		\multirow{2}{*}{Hardware} & Cluster Nodes                                                  & \begin{tabular}[c]{@{}l@{}}6 datanodes having \\ from 2 to 4 virtual \\ processors with 4 GB\\ RAM and namenode \\ having 8 GB RAM\end{tabular} & \begin{tabular}[c]{@{}l@{}}Intel Xeon @ 2Ghz\\ 8 cores, 16GB of RAM \\ and hyper threading\\ activated\end{tabular} \\ \cline{2-4}
-		& Networking                                                     & 1Gbit/second                                                                                                                                    &                                                                                                                     \\ \hline
+		\multirow{2}{*}{Hardware} & Cluster Nodes                                                  & \begin{tabular}[c]{@{}l@{}}6 data nodes with \\ from 2 to 4 virtual \\ processors and 4 GB\\ RAM and name node \\ with 8 GB RAM\end{tabular} & \begin{tabular}[c]{@{}l@{}}Intel Xeon @ 2Ghz\\ 8 cores, 16GB of RAM \\ and hyperthreading\\ activated\end{tabular} \\ \cline{2-4}
+		& Networking                                                     & 1 Gbit/second                                                                                                                                    &                                                                                                                     \\ \hline
 		\multirow{3}{*}{Software} & \begin{tabular}[c]{@{}l@{}}Java Virtual\\ Machine\end{tabular} & \begin{tabular}[c]{@{}l@{}}Java version \\ "1.7.0\_45"\\ Java SE Runtime\\ Environment\end{tabular}                                             & \begin{tabular}[c]{@{}l@{}}Java version \\ "1.7.0\_45"\\ Java SE Runtime\\ Environment\end{tabular}                 \\ \cline{2-4}
-		& Hadoop                                                         & \begin{tabular}[c]{@{}l@{}}hadoop 2.5\\ cloudera CDH5\end{tabular}                                                                              &                                                                                                                     \\ \cline{2-4}
+		& Hadoop                                                         & \begin{tabular}[c]{@{}l@{}}Hadoop 2.5\\ Cloudera CDH5\end{tabular}                                                                              &                                                                                                                     \\ \cline{2-4}
 		& \begin{tabular}[c]{@{}l@{}}Operating\\ System\end{tabular}     & CentOS 6.5                                                                                                                                      & CentOS 6.5                                                                                                          \\ \hline
@@
 
-The desktop implementation of the Algorithm is a java program for processing a small image dataset.
-As soon as the data set is processed the program halts.
-The desktop implementation starts as a single thread.
-One image is processed every time.
-As soon as the image is processed, it starts processing a new image.
-When the data set size is low, the program runs smooth with best performance.
-As data set size increases to Tera-bytes the system can't handle such huge amounts of data.
-To handle such large amounts of data we use the distributed computing using hadoop.
+The desktop implementations of the algorithms are Java programs for
+processing a small image dataset. As soon as the data set is processed
+the program halts. The desktop implementation starts as a single
+thread, and images are processed one at a time in succession.  When
+the dataset size is low, each program runs smoothly and with excellent
+performance.  However, as the size of the data to be processed reaches
+the terabyte scale, such a single-process approach fails utterly.
+Data on this scale must be processed in distributed fashion using
+Hadoop.
 
-The Algorithm is implemented on a hadoop cluster without using framework.
-It overcomes the problem of handling large datasets but suffers with a problem of code complexity.
-The user needs to write a huge amount of code and understand the technical details of hadoop.
-Writing such huge amounts of code including the custom InputFormats and RecordReaders is cumbersome.
-Fig.\ref{fig:module-chart} compares the actual lines of code user needs to write to run an Image Processing application in hadoop cluster which makes an average user life harder to use the hadoop features.
+Each algorithm is also implemented on a Hadoop cluster without using
+the Hadoop Image Processing Framework.  These implementations overcome
+the problem of handling large datasets but they suffer from extremely
+complex code.  The user needs to write a great deal of code and
+understand the technical details of Hadoop. Writing such code,
+including custom InputFormats and RecordReaders, is cumbersome and
+error-prone. Fig. \ref{fig:module-chart} compares the actual lines of
+code a user needs to write in order to run our experimental algorithms
+(or any similar image processing application) on a Hadoop cluster.
 
-The Algorithm is finally implemented on a hadoop cluster using the framework.
-It overcomes many problems handling large datasets, code complexity, interoperability between image datatypes etc., The process of creating CustomInputFormat and RecordReader is automated as per the input.
-User can tune the settings using different getter and setter method to get the desired output.
-The framework is so transparent that the user needs to input few lines of code to process a large image data set.
-It hides all the MapReduce Technical details and runs the code highly parallelized.
-The key differences among the three platform-specific implementation lie in computation elasticity and code complexity associated with each platform.
+Finally, we compare the same algorithms implemented on a Hadoop
+cluster using the framework.  These implementations overcome the
+problems of handling large datasets and managing interoperability
+between image datatypes while holding code complexity down to the same
+level as the single-process, local implementation.  Users enjoy
+automated creation of custom InputFormats and RecordReaders, and can
+tune settings for the desired output.  The framework is so transparent
+that the user needs only a few lines of code to process a large image
+data set. It hides the technical details of MapReduce and runs the
+code in highly parallelized fashion.  The code required to implement
+each image processing algorithm is no more complex than the same
+algorithm implemented on a local, single-threaded system, but the
+performance improvement for large datasets is immense.
 
@@
 		Desktop                                                                                             & \begin{tabular}[c]{@{}l@{}}Low : limited by \\ RAM and CPU of \\ the executing\\ computer\end{tabular} & \begin{tabular}[c]{@{}l@{}}Normal : Complexity \\ lies in writing the \\ algorithm\end{tabular}                                                                                                                                           & \begin{tabular}[c]{@{}l@{}}All data are \\ on local disk\end{tabular}                                                                                            \\ \hline
-		\begin{tabular}[c]{@{}l@{}}Hadoop:\\ without \\ using \\ framework\end{tabular}                     & \begin{tabular}[c]{@{}l@{}}High: Nodes can\\  be requested \\ as needed\end{tabular}                   & \begin{tabular}[c]{@{}l@{}}Heavy : Complexity \\ lies inwriting every \\ module in framework \\ including custom \\ Input Formats and \\ RecordReaders.User \\ should have high \\ technical knowledge on \\ hadoop working.\end{tabular} & \begin{tabular}[c]{@{}l@{}}All data resides\\ in HDFS.
-Custom\\ InputFormats \\ needsto be \\ created to launch\\ computations \\ where thedata are\end{tabular} \\ \hline
-		\begin{tabular}[c]{@{}l@{}}Hadoop:\\ using \\ framework\end{tabular}                                & \begin{tabular}[c]{@{}l@{}}High: Nodes can \\ be requested \\ as needed\end{tabular}                   & \begin{tabular}[c]{@{}l@{}}Normal : Complexity \\ lies in writing \\ algorithm and is \\ equivalent to writing \\ in Desktop platform\end{tabular}                                                                                        & \begin{tabular}[c]{@{}l@{}}All data resides \\ in HDFS.
+		\begin{tabular}[c]{@{}l@{}}Hadoop:\\ without \\ using \\ framework\end{tabular}                     & \begin{tabular}[c]{@{}l@{}}High: Nodes can\\  be requested \\ as needed\end{tabular}                   & \begin{tabular}[c]{@{}l@{}}Heavy : Complexity \\ lies in writing every \\ module in framework \\ including custom \\ InputFormats and \\ RecordReaders.
+User \\ requires high \\ technical expertise \\ with Hadoop.\end{tabular} & \begin{tabular}[c]{@{}l@{}}All data resides\\ in HDFS.
+Custom\\ InputFormats \\ need to be \\ created to launch\\ computations \\ in appropriate \\ locations\end{tabular} \\ \hline
+		\begin{tabular}[c]{@{}l@{}}Hadoop:\\ using \\ framework\end{tabular}                                & \begin{tabular}[c]{@{}l@{}}High: Nodes can \\ be requested \\ as needed\end{tabular}                   & \begin{tabular}[c]{@{}l@{}}Normal : Complexity \\ lies in writing \\ algorithm and is \\ equivalent to writing \\ on desktop platform\end{tabular}                                                                                        & \begin{tabular}[c]{@{}l@{}}All data resides \\ in HDFS.
 Highly \\ parallelized.\end{tabular}                                                                     \\ \hline
@@
 
-We have documented the following samples in the space of image processing
-computations, and hardware and software configurations applied to the selected
-data set.
+We have explored the space of image processing algorithms, applying
+different hardware and software configurations to the various data
+sets.  Specifically, we compared:
+
 \begin{itemize}
-	\item Three image processing computations described in Table 1.
-	\item Sampled data set of different image counts.
-	\item Hardware cluster configurations used for computations. 
-	\item Coding complexities in different environments for the three image processing computations
+\item Three image processing algorithms described in Section
+  \ref{algorithms}
+\item Datasets of varying image counts
+\item Configurations of hardware cluster used for computations 
+\item Coding complexity (measured in Lines of Code (LOC)
 \end{itemize}
 
-The numerical results are shown in Figure 1 through 6.
+The numerical results are shown in Figures \ref{fig:perf-chart}
+through \ref{fig:files-chart}.
 
-Figure \ref{fig:perf-chart} illustrates the performance comparison of Single Node, Hadoop using framework and Hadoop without using the framework for Canny Edge Detection.
+Figure \ref{fig:perf-chart} illustrates the performance comparison of
+Canny edge detection on a single node, on Hadoop using our framework
+and Hadoop without using the framework.
 
@@
   \includegraphics[width=0.45\textwidth]{perf-graph}
-	\caption{Canny Edge Detection computations executed on single node and on a Hadoop cluster with and without using framework}
+  \caption{Canny edge detection computation executed on a single node
+    and on a Hadoop cluster with and without using framework}
   \label{fig:perf-chart}
 \end{figure}
-It can be observed that the performance hadoop cluster is almost the same with and without using framework.
-The framework performs well sometimes because of multi-threading is implemented few tasks.
-Considering the single node performance it can be seen that the single node outperforms hadoop if the input is small as the input size increases handling huge data becomes infeasible.
+
+It can be observed that the Hadoop cluster performance is nearly
+identical whether the Hadoop Image Processing Framework is used or
+not.  The single node performance is satisfactory on small input
+sizes; in fact it outperforms Hadoop when the amount of data is very
+small owing to the lack of MapReduce overhead.  However, as the amount
+of data grows the single node performance becomes increasingly poor.
 The same pattern is observed with any image processing algorithm.
 
-Figure \ref{fig:module-chart} compares coding complexities of different modules the framework described.
-The Downloader module and Extractor module are simple to use and require less number of lines of code to configure.
-The processor module includes the algorithm the user devised, which adds up a few extra lines of code.
-but configuring the devised algorithm to a MapReduce job is fairly simple with very less number of steps.
-The modules are included with different methods for flexibility.
+Figure \ref{fig:module-chart} compares the coding complexity of the
+different image handling and processing tasks described in Section
+\ref{methodology}, for the Canny edge detection task. The Downloader
+module and Extractor module are simple to use and require few lines of
+code to configure. The processor module includes the specific
+algorithm devised by the user, which can be significant amount of
+code, but using our framework to configure the algorithm for MapReduce
+requires almost no additional coding.
 
@@
   \includegraphics[width=0.45\textwidth]{module-chart2}
-	\caption{Comparing coding complexities of modules in different environments}
+  \caption{Comparing coding complexities of modules in different
+    environments}
   \label{fig:module-chart}
@@
 
-Figure \ref{fig:comp-chart} shows the coding complexities of three image processing computations from Table 1.
-Single node and the framework show similar coding complexity outperforming the image processing without framework.
-The reason behind this is Hadoop Image Processing Framework hides the inner details of MapReduce programming which makes an user to think that he is working on the single node but actually the task is processed in hadoop.
-Without the framework user needs to write a MapReduce job which is cumbersome.
+Figure \ref{fig:comp-chart} shows the coding complexity of the
+processor portion of our three experimental image processing
+algorithms.  The naive single-process, single-node algorithm requires
+a similar amount of code to using our highly parallel, distributed
+framework, while direct implementation of the algorithm using Hadoop
+and MapReduce requires a great deal more effort.  The Hadoop Image
+Processing Framework hides all of this complexity from the user.
 
@@
   \includegraphics[width=0.45\textwidth]{comp-chart}
-	\caption{Comparing coding complexities of laplacian filter, canny edge detection and k-means clustering  in different environments}
+  \caption{Comparing coding complexity of Laplacian filter,
+    Canny edge detection and $k$-means clustering in different
+    environments}
   \label{fig:comp-chart}
@@
 
-Hadoop Image processing framework provides transparency over two levels.
+The Hadoop Image processing framework provides transparency over two levels.
 The first level of transparency allows user to use the predefined MapReduce modules Downloader, Processor and Extractor for processing the image modules.

### 2015-03-15 Paper.tex
Updated footnotes and references

@@
 storage.  The emergence of processing frameworks such as the Hadoop
-MapReduce\cite{Hadoop2009} platform addresses the problem of providing
+MapReduce\cite{dean2008} platform addresses the problem of providing
 a system for computationally intensive data processing and distributed
@@
 applications that can use large sets of images has been limited. To
-address this we have developed the Hadoop Image Processing Framework,
+address this we have developed the Hadoop Image Processing Framework\footnote[1]{S. Vemula, C. Crick Hadoop Image Processing Framework https://github.com/okstate-robotics/hipl.git},
 which provides a Hadoop-based library to support large-scale image
@@
 \subsection{Downloading and storing image data}
-Hadoop uses the Hadoop Distributed File System (HDFS) to store files
+Hadoop uses the Hadoop Distributed File System (HDFS)\cite{Shvachko2010} to store files
 in various nodes throughout the cluster.  One of Hadoop's significant

### 2015-03-15 Paper.tex
Everything but conclusion and related work.

--- 2015-03-15 Paper.tex
@@
 
-The Hadoop Image processing framework provides transparency over two levels.
-The first level of transparency allows user to use the predefined MapReduce modules Downloader, Processor and Extractor for processing the image modules.
-The second of level of transparency allows the users to write custom map reduce programs as the framework is built in an hierarchal manner.
+The Hadoop Image processing framework provides transparency on
+multiple levels.  Users can use the predefined MapReduce modules
+(Downloader, Processor and Extractor) for processing the image
+modules, and most image processing tasks require no more than this.
+Users may also write custom MapReduce tasks within the framework,
+taking advantage of its hierarchical construction.  This still
+protects the user from the full complexity and error-prone nature of
+the full Hadoop framework, but requires additional knowledge and
+expertise on the user's part.
 
-The great thing about Hadoop Image Processing framework is its built in such a way that using it is really simple.
-The framework strictly adheres java file writing and reading techniques and extends it to a new level of BundleFile Writing and Reading using hadoop framework.
-This makes user to work with hadoop image processing more flexible and easier.
-The main aim of the framework is to provide an abstractness in such a way that programming on a single computer is equivalent to programming on a hadoop cluster.
-Listing 1.
-provides the similarities in code writing in java and image processing framework.
-Making framework similar to java helps understanding the framework more easier and reduces complexity.
-Listing 2 shows a sample code for setting image headers for processed images.
+The Hadoop Image Processing Framework is intended to be extremely
+simple to use. The framework strictly adheres to Java file writing and
+reading techniques, extending those conventions to the Hadoop
+framework's notion of reading and writing bundle files.  In this way,
+working with Hadoop image processing is exactly like working on a
+single system.  The main aim of the framework is to provide useful
+software abstractions such that programming on a Hadoop cluster is
+equivalent to programming on a single computer.  Listing 1
+demonstrates the similarity between Java code written for a single
+machine and the same code using the image processing framework.  The
+framework's use of ordinary Java conventions helps in understanding
+the software engineering process and reduces complexity.  Listing 2
+demonstrates the mechanism for setting the image headers of processed
+images.
 
@@
   \includegraphics[width=0.45\textwidth]{files-chart2}
-	\caption{Comparing File Writers and readers without and with using framework }
+  \caption{Comparing file writers and readers with and without using
+    framework}
   \label{fig:files-chart}

### 2015-03-15 Paper.tex
Finished editing, except for related work.

@@
 \section{Conclusion}
-This paper has described our framework fo image processing for large scale image processing applications on the MapReduce framework.
-The framework is designed to abstract the technical details of Hadoop's powerful MapReduce framework and provide a way for users who want to process large image datasets.
-We provided a how the images can be stored in different hadoop file format and efficient access within Map Reduce pipeline and simple modules to make the framework more easier.
-By providing interoperability between different image data types we give the user to process the images using different open-source image processing libraries.
-Finally, we provide a way to overcome the loss of image headers after processing which are most useful for image processing and vision applications.
+This paper has described our Hadoop Image Processing Framework for
+implementing large scale image processing applications.  The framework
+is designed to abstract the technical details of Hadoop's powerful
+MapReduce system and provide an easy mechanism for users to process
+large image datasets. We provide software machinery for storing images
+in the various Hadoop file formats and efficiently accessing the Map
+Reduce pipeline.  By providing interoperability between different
+image data types we allow the user to leverage many different
+open-source image processing libraries. Finally, we provide the means
+to preserve image headers throughout image manipulation process,
+retaining useful and valuable information for image processing and
+vision applications.
 
-With these features, the framework brings a new level of transparency and simplicity for creating large-scale image processing applications over Hadoop's MapReduce framework.
-The paper describes to example applications built using the framework to demonstrate the power of the framework.
-We hope that this level of abstractness to use Hadoop MapReduce framework will enhance the ability to create large-scale image processing applications with ease.
+With these features, the framework provides a new level of
+transparency and simplicity for creating large-scale image processing
+applications on top of Hadoop's MapReduce framework.  We demonstrate
+the power and effectiveness of the framework in terms of performance
+enhancement and complexity reduction.  The Hadoop Image Processing
+Framework should greatly expand the population of software developers
+and researchers easily able to create large-scale image processing
+applications.
 
 \section{Future Work}
-In the near future we hope to extend the framework into a full-fledged multimedia processing framework.
-We would like to improve framework such that audio and video processing over hadoop is done with ease.
-We also intend to add a CUDA module for using the graphics card for processing the images.
-Finally, we intend to wrap our system into an highly parallelized open-source hadoop multimedia processing framework and perhaps provide a sample web-based graphic user interface with all image processing applications.
+In the near future we hope to extend the framework into a full-fledged
+multimedia processing framework. We would like to improve the
+framework to handle audio and video processing over Hadoop with
+similar ease. We also intend to add a CUDA module to allow processing
+tasks to make use of machines' graphics cards. Finally, we intend to
+develop our system into a highly parallelized open-source Hadoop
+multimedia processing framework, providing web-based graphical user
+interfaces for image processing applications.
 

### 2015-03-15 Paper.tex
Possibly final edited version.

@@
 applications that can use large sets of images has been limited. To
-address this we have developed the Hadoop Image Processing Framework\footnote[1]{S. Vemula, C. Crick Hadoop Image Processing Framework https://github.com/okstate-robotics/hipl.git},
+address this we have developed the Hadoop Image Processing Framework,
 which provides a Hadoop-based library to support large-scale image
@@
 advantage of the huge data stores, rich metadata and global reach of
-current online image sources.
+current online image sources.\footnote[1]{The Hadoop Image Processing
+  Framework is open source and freely available for download at
+  https://github.com/okstate-robotics/hipl.git}.
 
@@
 available than ever before, and is growing every day.  Every minute
-27800 photos are uploaded to Instagram, \cite{Horaczek2013}while Facebook receives
-208,300 photos over the same time frame. This alone
+27,800 photos are uploaded to Instagram,\cite{Horaczek2013} while
+Facebook receives 208,300 photos over the same time frame. This alone
 provides a source of image data that can scale into the billions.  The
@@
 
-White et.al \cite{White2010} presents a case study of classifying and clustering billions of regular images using MapReduce.
-It describes an image pre-processing technique for use in a sliding-window approach for object recognition.
-Pereira et.al \cite{Pereira2010} outlines some of the limitations of the MapReduce model when dealing with high-speed video encoding, namely its dependence on the NameNode as a single point of failure, and lack of possibility for generalization in order to suit the issue at hand.
-An alternate optimized implementation is proposed for providing cloud-based IaaS (Infrastructure as a Service) solution.
-Zhenhua Lv et.al \cite{Lv2010} describe using the k-means algorithm in conjunction with MapReduce and satellite/aerophoto images in order to find different elements based on their color.
+White et.al \cite{White2010} presents a case study of classifying and
+clustering billions of regular images using MapReduce.  It describes
+an image pre-processing technique for use in a sliding-window approach
+for object recognition.  Pereira et.al \cite{Pereira2010} outlines
+some of the limitations of the MapReduce model when dealing with
+high-speed video encoding, namely its dependence on the NameNode as a
+single point of failure, and the difficulties inherent in generalizing
+the framework to suit particular issues.  It proposes an alternate
+optimized implementation for providing cloud-based IaaS
+(Infrastructure as a Service) solutions.  Lv et.al \cite{Lv2010}
+describes using the $k$-means algorithm in conjunction with MapReduce
+and satellite/aerial photographs in order to find different elements
+based on their color.
 
-Zhang et.al \cite{Zhang2010} present methods used for processing sequences of microscope images of live cells.
-The images are relatively small 512x512 16-bit pixels, stored in 90 MB folders there were some issues regarding to fitting into Hadoop DFS blocks with were solved by custom InputFormat, InputSplit and RecordReader classes.
-Mark Powell et.al \cite{Powell2010} describes the way NASA handles image processing of celestial images captured by the Mars Orbiter and rovers.
-Clear and concise descriptions are provided about the segmentation of gigapixel images into tiles, how the tiles are processed and how image processing framework handles scaling and works with the distributed processing.
-Wang, Yinhai and McCleary\cite{Wang2011} about the speeding up the analysis of Tissue MicroArray images by substituting human expert analysis for automated processing algorithms.
-While the images were measured to be in gigapixels, the content is easily segmented and there was no need to focus on being able to analyze all of image at once.
-The work was all done on a specially build grip high performance	computing platform on Hadoop framework.
+Zhang et.al \cite{Zhang2010} presents methods used for processing
+sequences of microscope images of live cells. The images are
+relatively small (512x512, 16-bit pixels).  Stored in 90 MB folders,
+the authors encountered difficulties regarding fitting into Hadoop DFS
+blocks with were solved by custom InputFormat, InputSplit and
+RecordReader classes.  Powell et.al \cite{Powell2010} describes how
+NASA handles image processing of celestial images captured by the Mars
+Orbiter and rovers. Clear and concise descriptions are provided about
+the segmentation of gigapixel images into tiles, how the tiles are
+processed and how the image processing framework handles scaling and
+works with the distributed processing. Wang, Yinhai and
+McCleary\cite{Wang2011} discuss speeding up the analysis of tissue
+microarray images by substituting human expert analysis for automated
+processing algorithms. While the images were gigapixel-sized, the
+content was easily segmented and there was no need to analyze all of
+an image at once. The work was all done on a specially-built high
+performance computing platform using the Hadoop framework.
 
-Bajcsy et.al \cite{Bajcsy2013} present a characterization of four basic terabyte size image computations on a Hadoop Cluster in terms of their relative efficiency according to the modified Amdahls law.
-The work is motivated by the fact that there is a lack of 6 standard benchmarks and stress tests for big image processing operations on Hadoop framework.
-Moise et.al \cite{Moise2013} outlines about querying thousands of images in one run using Hadoop MapReduce framework using the MapReduced based eCP Algorithm.
-The experiment does an image search on 110 million images collected from web on Grid 5000 Platform.
-The results are evaluated in order to understand the best practices that can tune Hadoop MapReduce Performance for image search.
+Bajcsy et.al \cite{Bajcsy2013} present a characterization of four
+basic terabyte-size image computations on a Hadoop cluster in terms of
+their relative efficiency according to a modified Amdahl's Law. The
+work was motivated by the fact that there is a lack of standard
+benchmarks and stress tests for large-scale image processing
+operations on the Hadoop framework. Moise et.al \cite{Moise2013}
+outlines the querying of thousands of images in one run using the
+Hadoop MapReduce framework and the eCP Algorithm. The experiment
+performs an image search on 110 million images collected from the web
+using the Grid 5000 platform. The results are evaluated in order to
+understand the best practices for tuning Hadoop MapReduce performance
+for image search.
 
-While the above shows that there has been lot of work in this area, providing a framework to ease the process develop large-scale image processing applications.
-The question remains how the performance for Hadoop can be increased for large scale image processing tasks on small files.
+The above shows that there has been a great deal of intensive work in
+image processing using MapReduce.  However, each independent project
+requires a complex, error-prone, one-off implementation.  Such
+research and application development would benefit greatly from a
+standard, well-engineered image processing framework such as the one
+we provide.
 

### 2015-03-15 Paper.tex
Fixed spacing issue.
minor text changes and table descriptions

@@
 image datasets are made up of small image files in great numbers,
-which is a situation which HDFS has a great deal of trouble
-handling. The problem can be solved by providing a container to group
+which is a situation HDFS has a great deal of trouble
+handling. This problem can be solved by providing a container to group
 the files in some way. Hadoop offers a few options:
@@
 extracted from a database or a provided by a search. The framework
-provides a extendable ripper module for extracting URLS from Flickr
+provides an extendable ripper module for extracting URLs from Flickr
 and Google image searches and from SQL databases. In addition to the
@@
 across the available processing nodes for maximum efficiency and
-parallelism.  Each node map task generates several image bundles
-appropriate to the selected input type, containing all of the images
-it downloaded. In the reduce phase, the Reducer will merge these image
+parallelism. The URL list is split into several map tasks of equal size across the nodes. Each node map task generates several image bundles
+appropriate to the selected input list, containing all of the image URLs
+to download. In the reduce phase, the Reducer will merge these image
 bundles into a large image bundle.
@@
 \subsection{Processing image bundle using MapReduce}
-Hadoop MapReduce programs handle input and output data very
+Hadoop MapReduce program handles input and output data very
 efficiently, but their native data exchange formats are not convenient
@@
 Hadoop Image Processing Framework provides such InputFormat and
-RecordReaders for the sake of the system's ease of use.
+RecordReaders for system's ease of use.
 
@@
 results.  The small amount of storage overhead required for this
-functionality is a trade-off worth making for the sake of preserving
+functionality is a trade-off worth making for preserving
 image header data after processing.
@@
 individual map tasks. Each map task will apply the processing
-algorithm to each local image and append them in turnto the output
+algorithm to each local image and append them to the output
 image bundle. In the reduce phase, the Reducer merges these image
@@
 
-\textbf{Step 2: Split Image bundle across nodes} The input image
+\textbf{Step 2: Split Image bundle across nodes.} The input image
 bundle is split across available nodes using the framework's custom
@@
 
-\textbf{Step 3: Extract individual image} The image bytes obtained
+\textbf{Step 3: Extract individual image.} The image bytes obtained
 from each HImage object are stored onto the filesystem in the
@@
 detection and $k$-means image segmentation.  These are widely-used,
-computation- and data-intensive algorithms of varying complexity which
+computation-intensive and data-intensive algorithms of varying complexity which
 require large distributed systems for timely operation on hundreds of
@@
 benefits of the Hadoop system while retaining the simplicity of the
-single-node implementation.  In this section we describe the dataset
-and hardware and software specifications used in our experiments, and
+single-node implementation.  In this section we describe the dataset, 
+hardware and software specifications used in our experiments, and
 discuss the results.
@@
 \subsection{Characteristics of Image Dataset, Hardware and Software}
-We performed image processing computations on a dataset of about 1TB
+We performed image processing computations on a dataset of about 1 TB
 in size. The dataset was obtained by performing a Flickr search based
@@
 Hadoop Image Processing Framework.  The framework currently provides
-rippers for Flickr and Google Image searches, and more will be
+rippers for Flickr and Google image searches, and more will be
 provided in the future.  A Ripper interface extracts image URLs from
 search results, and can be implemented for any set of results in any
-format.  The experimental dataset is composed of 220000 images at 4.76
+format.  The experimental dataset is composed of 220,000 images at 4.76
 MB/image, or about 1 TB.
 
-\subsection{Computer Hardware and Software Characterstics}
+\subsection{Computer hardware and software characteristics}
 We ran the image processing computations on a 6-node distributed
@@
 \begin{table}[h]
+	\caption {Computer hardware and software characteristics.}
+	\label{table:table1}
 	\begin{tabular}{|l|l|l|l|}
@@
 	\end{tabular}
+
 \end{table}
@@
 including custom InputFormats and RecordReaders, is cumbersome and
-error-prone. Fig. \ref{fig:module-chart} compares the actual lines of
+error-prone. Figure \ref{fig:module-chart} compares the actual lines of
 code a user needs to write in order to run our experimental algorithms
@@
 \begin{table}[h]
+	\caption{Characteristics of computational elasticity, coding complexity and data locality for the three experimental platforms.}
+	\label{table:table2}
 	\begin{tabular}{|l|l|l|l|}
@@
 \item Configurations of hardware cluster used for computations 
-\item Coding complexity (measured in Lines of Code (LOC)
+\item Coding complexity - measured in Lines of Code (LOC)
 \end{itemize}
@@
 
-The Hadoop Image processing framework provides transparency on
+The Hadoop Image Processing Framework provides transparency on
 multiple levels.  Users can use the predefined MapReduce modules
@@
 
-\begin{lstlisting}[caption = Comparing File Writer instance in java and SequenceBundleWriter instance in Hadoop Image Processing framework ]
-
+\begin{lstlisting}[caption = Comparing FileWriter instance in java and SequenceBundleWriter instance in Hadoop Image Processing framework ]
 //Sample File Writing in Java
 
-File file = new File("abc.txt");
-FileWriter fw=new FileWriter(file);
+File file = new File("test");
+FileWriter fw = new FileWriter(file);
 fw.append(val)
 fw.close();
-
 //Sample BundleFile Writing in Hadoop using Framework
 
-BundleFile file = new BundleFile("abc.seq");
-SequenceBundleWriter sbw=new SequenceBundleWriter(file);
+BundleFile file = new BundleFile("test_bundle.seq");
+SequenceBundleWriter sbw = new SequenceBundleWriter(file);
 sbw.append(himage);
 sbw.close();
-
 \end{lstlisting}

### 2015-03-16 Paper.tex
Fixed a widow.

@@
 \end{figure}
+\newpage
+\begin{lstlisting}[caption = Comparing FileWriter instance in java and SequenceBundleWriter instance in Hadoop Image Processing framework ]
 
-\begin{lstlisting}[caption = Comparing FileWriter instance in java and SequenceBundleWriter instance in Hadoop Image Processing framework ]
 //Sample File Writing in Java
-
 File file = new File("test");
@@
 fw.close();
+
 //Sample BundleFile Writing in Hadoop using Framework
-
 BundleFile file = new BundleFile("test_bundle.seq");
@@
 //input - HImage is sent as input 
-
 CannyEdge canny = new CannyEdge(input);
